# Executive summary {.unnumbered}
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

county_emissions_summary <- county_emissions %>%
  group_by(sector, category, source, data_source, factor_source) %>%
  summarize(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e),
    .groups = "keep"
  )

hookaddcap()
```

*This project has been funded wholly or in part by the United States Environmental Protection Agency (EPA) under assistance agreement 00E03476 to the Metropolitan Council. The contents of this document do not necessarily reflect the views and policies of the EPA, nor does the EPA endorse trade names or recommend the use of commercial products mentioned in this document.*

::: {.callout-note}
This document is one part of the Metropolitan Council Priority Climate Action Plan (PCAP), submitted to the EPA on March 1, 2024. See [metrocouncil.org](https://www.metrocouncil.org/tcghginventory) for more details. 
:::

```{r fig-ghg-inventory-tableau}
#| fig-cap: "Greenhouse Gas Inventory Interactive "
#| out-width: "95%"
#| out-height: "800px"
# htmltools::HTML("<iframe width='95%' height='500' src='https://metrocouncil.maps.arcgis.com/apps/webappviewer/index.html?id=0b0735b3407f49ceb347fc30c9b83bda' title='Regional Bicycle Transportation Network'></iframe>")
```

```{r sector-percentages, echo=FALSE, include=FALSE}
sector_percentages <- county_emissions_summary %>%
  group_by(sector) %>%
  summarize(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)) %>%
  arrange(-emissions_metric_tons_co2e) %>%
  janitor::adorn_percentages(denominator = "col")
sector_percentages
```

In 2021, the Twin-Cities MSA generated `r round(sum(county_emissions$emissions_metric_tons_co2e)/1000000, digits = 2)` million metric tons CO~2~ equivalent (MMtCO~2~e) emissions from energy, transportation, and waste sectors. `r sector_percentages[[1]][1]` was the largest contributor to GHG emissions (`r scales::percent(sector_percentages[[2]][1], accuracy = 0.1)`), followed by `r sector_percentages[[1]][2]` (`r scales::percent(sector_percentages[[2]][2], accuracy = 0.1)`), and `r sector_percentages[[1]][3]` (`r scales::percent(sector_percentages[[2]][3], accuracy = 0.1)`). 

```{r fig-county-emissions-sector}
#| fig-cap: "County emissions by sector and category"
#| out-width: "95%"

fig_county_emissions_sector <- plot_ly(
  data = county_emissions %>%
    group_by(sector, category) %>%
    summarize(
      emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e),
      .groups = "keep"
    ),
  source = "fig-county-emissions-sector",
  x = ~sector,
  y = ~emissions_metric_tons_co2e,
  color = ~category,
  colors = unlist(category_colors),
  split = ~category,
  type = "bar",
  hovertemplate = ~ paste0(
    category, "<br>",
    round(emissions_metric_tons_co2e * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout("11-county regional greenhouse gas emissions, 2021",
    subtitle = "",
    x_title = "Sector",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Category",
    barmode = "stack"
  )

fig_county_emissions_sector
```

## Emissions by sector

The energy sector generated `r round(sum(county_emissions$emissions_metric_tons_co2e)/1000000 * sector_percentages[[2]][1], digits = 2)` MMtCO~2~e of emissions in the Twin Cities MSA in 2021. Most of the energy emissions are from electricity followed by natural gas. Other fuels (propane and kerosene) make up a small proportion of energy emissions in the metro.


```{r fig-energy-emissions}
#| fig-cap: "Energy county level emissions"
#| out-width: "95%"
fig_energy_emissions <- plot_county_emissions(
  county_emissions = county_emissions,
  .sector = "Energy",
  .plotly_source = "fig-energy-emissions"
)
fig_energy_emissions
```

The transportation sector generated `r round(sum(county_emissions$emissions_metric_tons_co2e)/1000000 * sector_percentages[[2]][2], digits = 2)` MMtCO~2~e of emissions in the Twin Cities MSA in 2021.

```{r fig-transportation-emissions}
#| fig-cap: "Transportation county level emissions"
#| out-width: "95%"
fig_transportation_emissions <- plot_county_emissions(
  county_emissions = county_emissions,
  .sector = "Transportation",
  .plotly_source = "fig-transportation-emissions"
)
fig_transportation_emissions
```

Emissions from waste is the smallest of the three sectors in the Twin Cities MSA. Waste generated `r round(sum(county_emissions$emissions_metric_tons_co2e)/1000000 * sector_percentages[[2]][3], digits = 2)` MMtCO~2~e of emissions in the Twin Cities MSA in 2021. Solid waste, including landfills, recycling, and organics, generates the largest share of emissions in the waste sector, with wastewater treatment comprising a smaller share of waste emissions. 

```{r fig-waste-emissions-index}
#| fig-cap: "Waste and wastewater county level emissions"
#| out-width: "95%"
fig_waste_emissions_index <- plot_county_emissions(
  county_emissions = county_emissions,
  .sector = "Waste",
  .plotly_source = "fig-waste-emissions-index"
)

fig_waste_emissions_index$x$layoutAttrs[[1]]$title$text <- "Solid waste and wastewater emissions"

fig_waste_emissions_index
```

::: {.content-visible when-format="html"}

## About this document

This is a Quarto book. Code, data, and documentation are all stored in the same place. We primarily use R. 

- To navigate from chapter to chapter, click on the chapter name in the left sidebar or click the arrow button at the end of the page.
- Citations can be previewed by hovering your mouse over the citation.
- Links to R scripts are available throughout the document. Click on the link to open the relevant R script and peruse the code behind this inventory.
- All plots are interactive. Hover your mouse over any bar or dot, and detailed information will appear. If you want to take a static screenshot of a plot, hover your mouse over the plot and find the camera icon in the upper right tool bar.
- Sections are cross-referenced throughout the document. Click on a "Section x.xx.xx" link to navigate to the appropriate section
- All headings, tables, and figures are permalinked, meaning that you can share specific sections of the document through a URL. Hover your mouse over a heading and click on the link icon. Then, copy the URL in your browser and share with colleagues. 

::: 

## Acknowledgements

This document is the result of tremendous work across the Metropolitan Council. Individuals are noted by their name, title, contribution, and division. 

:::: {.columns}

::: {.column width="50%"}
#### Primary authors {.unnumbered}
 - Liz Roten, Senior Data Scientist, Transportation (MTS)  
 - Sam Limerick, Senior Data Scientist, Electricity and natural gas (CD)  
 - Peter Wilfahrt, Principal Researcher, Liquid stationary energy and wastewater (CD)   
 - Zoey Yandell, Intern, Solid waste (CD)  
:::

::: {.column width="50%"}
#### Contributors {.unnumbered}
 - Kenneth Smith, Principal Researcher, LIDAC, editing (CD)  
 - Laine McNamara, Business Systems Analyst III, editing (CD)  
 - Trevor Prater, Senior Engineer, wastewater (ES)  
:::

::::

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```

{{< pagebreak >}}

