# Supplementary data
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
hookaddcap()
```

## Geographic data

Geographic data were processed in [cprg_geography.R](../R/cprg_geography.R).

### Counties

County data was pulled using [`{tigris}`](@walkerTigrisLoadCensus2023), an R package that downloads TIGER/Line shapefiles from the US Census Bureau. 2021 geographies for Minnesota and Wisconsin were pulled, combined, and saved. 

```{r tbl-cprg-county}
#| tbl-cap: "County geography metadata"
#| out-width: "95%"
cprg_county_meta <- readRDS(file.path(here::here(), "R/data/cprg_county_meta.RDS"))

cprg_county_meta %>% 
  kable()
```


### Cities

Minnesota cities, townships, and unorganized territories were imported from [Minnesota Geospatial Commons](https://gisdata.mn.gov/dataset/bdry-mn-city-township-unorg) @mndotCityTownshipUnorganized2023. This data was then filtered on the `COUNTY_NAM` column to include only the workplan area counties.

Wisconsin cities, towns, and villages were imported from Wisconsin's Legislative Technology Services Bureau @wisconsinlegislatureWisconsinCitiesTowns2023. Data was then filtered on `CNTY_NAME` to include only the workplan area counties.

```{r tbl-cprg-ctu}
#| tbl-cap: "City geography metadata"
#| out-width: "95%"
cprg_ctu_meta <- readRDS(file.path(here::here(), "R/data/cprg_ctu_meta.RDS"))

cprg_ctu_meta %>% 
  kable()
```


```{r echo=FALSE}
saveCap("supp-data")
```
