# Agriculture {#sec-agriculture}
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))
ag_emissions <- readRDS(file.path(here::here(), "_agriculture/data/agricultural_emissions_county.RDS"))
cprg_county <- readRDS(file.path(here::here(), "_meta/data/cprg_county.RDS"))

emissions_2022 <- county_emissions %>%
  filter(emissions_year == 2022) %>%
  pull(value_emissions) %>%
  sum()
ag_emissions_2022 <- county_emissions %>%
  filter(emissions_year == 2022, sector == "Agriculture") %>%
  pull(value_emissions) %>%
  sum()
ag_emissions_2005 <- county_emissions %>%
  filter(emissions_year == 2005, sector == "Agriculture") %>%
  pull(value_emissions) %>%
  sum()


rural_ag_2022 <- county_emissions %>%
  filter(
    emissions_year == 2022, sector == "Agriculture",
    county_name %in% c("St. Croix", "Pierce", "Carver", "Dakota")
  ) %>%
  pull(value_emissions) %>%
  sum()

methane_2022 <- ag_emissions %>%
  filter(
    inventory_year == 2022,
    units_emissions == "Metric tons CH4"
  ) %>%
  pull(mt_co2e) %>%
  sum()

n2o_2022 <- ag_emissions %>%
  filter(
    inventory_year == 2022,
    units_emissions == "Metric tons N2O"
  ) %>%
  pull(mt_co2e) %>%
  sum()

hookaddcap()
```

## Introduction

Agricultural systems are major emitters of methane and nitrous oxide, both from livestock and croplands. Livestock emissions come primarily from enteric fermentation, the formation of methane in ruminants stomachs during digestion, and secondarily from manure emissions of both methane and nitrous oxides. Cropland emissions are predominantly nitrous oxide that is formed from biogeochemical processes related to crop-soil interactions and the direct application of synthetic fertilizers. Carbon dioxide emissions are a smaller share of agricultural emissions, arising predominantly from operation of agricultural equipment, liming of soils, and production of urea used for fertilizer.


## Results

### 2022 county and subsector breakdown

Agriculture accounted for `r 100*round(ag_emissions_2022/emissions_2022, digits = 3)`% of total emissions in the 11-county region in 2022. As expected, more rural counties produce a larger share of the agricultural emissions, with St. Croix, Pierce, Dakota, and Carver accounting for `r 100*round(rural_ag_2022/ag_emissions_2022, digits = 3)`% of the regions agricultural emissions.

```{r fig-ag-county-emissions}
#| fig-cap: "2022 county agricultural emissions"
#| out-width: "95%"
#| fig.height: 4  # Set height in inches
#| fig.width: 10   # Set width in inches

fig_ag_county_emissions <- plot_ly(
  type = "bar",
  source = "fig-ag-county-emissions",
  data = county_emissions %>%
    filter(
      sector == "Agriculture",
      emissions_year == 2022
    ),
  y = ~ reorder(county_name, value_emissions),
  x = ~value_emissions,
  color = ~source,
  colors = unlist(source_colors),
  hovertemplate = ~ paste0(
    county_name, "<br>",
    category, "<br>",
    source, "<br>",
    round(value_emissions / 1000000, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = "2022 county agricultural emissions",
    x_title = "Million metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_ag_county_emissions
```


### Baseline emissions

Regional agricultural emissions have decreased by `r 100*round(1 - (ag_emissions_2022/ag_emissions_2005),2)`% since 2005, potentially reflecting increased agricultural abandonment and suburban expansion in the region.

```{r fig-ag-baseline-emissions}
#| fig-cap: "Baseline agricultural emissions"
#| out-width: "95%"
#| fig.height: 4  # Set height in inches
#| fig.width: 8   # Set width in inches

fig_ag_baseline <- plot_ly(
  source = "fig-ag-baseline-emissions",
  data = county_emissions %>%
    filter(sector == "Agriculture",
           emissions_year <= 2022) %>%
    group_by(emissions_year, category) %>%
    summarise(value_emissions = sum(value_emissions)),
  x = ~emissions_year,
  y = ~value_emissions,
  color = ~category, #
  colors = unlist(category_colors), # Directly map the colors to categories
  type = "scatter",
  mode = "lines", # No points or lines, just fill
  stackgroup = "one", # Fill to the next y value (stacking)
  hoverinfo = "x+y+name", # Ensure hover displays x, y, and category name
  hovertemplate = ~ paste0(
    "Year: %{x}<br>",
    "Category: %{text}<br>",
    "Emissions: %{y:.2f} metric tons CO<sub>2</sub>e<br>",
    "<extra></extra>"
  ),
  text = ~category # Pass category to hovertext
) %>%
  plotly_layout(
    main_title = "County Agricultural Emissions Over Time",
    x_title = "Year",
    y_title = "Million metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    hovermode = "x unified", # Combine all hovertext under one hover
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_ag_baseline
```

### Emissions by gas type

The majority of agricultural emissions by weight are methane, but when accounting for the increased warming potential of nitrous oxide, CO<sub>2</sub>e the two are roughly equivalent. 
CH<sub>4</sub>: `r 100 * round(methane_2022/ag_emissions_2022, digits = 3)`% 
N<sub>2</sub>O: `r 100 * round(n2o_2022/ag_emissions_2022, digits = 3)`%


```{r fig-ag-county-emissions-by-gas-type-co2e}
#| fig-cap: "2022 county agricultural emissions by gas type (CO2e)"
#| out-width: "95%"

fig_ag_county_emissions_by_gas_type_co2e <- plot_ly(
  type = "bar",
  source = "fig-ag-county-emissions-by-gas-type-co2e",
  data = ag_emissions %>%
    left_join(., cprg_county) %>%
    filter(inventory_year == 2022),
  y = ~ reorder(county_name, mt_co2e),
  x = ~mt_co2e,
  color = ~units_emissions,
  hovertemplate = ~ paste0(
    county_name, "<br>",
    source, "<br>",
    round(mt_co2e, digits = 2), "  metric tons of CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = "2022 county agricultural emissions by gas type",
    x_title = "Metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_ag_county_emissions_by_gas_type_co2e
```

## CTU Estimates

Agricultural emissions are apportioned to cities and townships in two ways. First, county crop production and fertilizer application are apportioned to cities and townships by their percentage of the counties cropland acreage. Second, livestock are assumed to have minimal occurrences in city boundaries, so the headcounts are apportioned to cities and townships designated as Rural Diversified, Rural Residential, or Agricultural. Based on these activity data, we follow the same emission calculations as for counties.

```{r fig-ctu-ag-emissions}
#| fig-cap: "2022 city level agricultural emissions"
#| out-width: "95%"

# load the county boundaries layer

### load emissions and downscaling var (proportion of county ag land)
ctu_ag_emissions <- readRDS(file.path(here::here(), "./_agriculture/data/agricultural_emissions_ctu.rds"))

agricultural_emissions_2022 <- ctu_ag_emissions %>%
  filter(inventory_year == 2022) %>%
  group_by(ctu_name, ctu_class) %>%
  summarize(mt_co2e = sum(mt_co2e, na.rm = TRUE))

### map CTU scale agricultural emissions

cprg_ctu <- readRDS(here("_meta", "data", "cprg_ctu.RDS")) %>%
  mutate(geometry = st_make_valid(geometry))


ctu_ag_map <- left_join(cprg_ctu,
  agricultural_emissions_2022,
  by = c("ctu_name", "ctu_class")
) %>%
  select(ctu_name, geometry, mt_co2e) %>%
  replace(is.na(.), 0)


# Define color palette based on the plasma scale from viridis
pal <- colorNumeric(palette = "plasma", domain = ctu_ag_map$mt_co2e)

# Create the leaflet map
fig_ctu_ag_emissions <- leaflet(ctu_ag_map) %>%
  addTiles() %>% # Add default OpenStreetMap tiles
  addPolygons(
    fillColor = ~ pal(mt_co2e),
    color = "black", # Border color
    weight = 1,
    fillOpacity = 0.9, # Transparency of the fill color
    popup = ~ paste(ctu_name, " agricultural emissions:", round(mt_co2e, 2)), # Popup with details
    label = ~ paste(ctu_name, " emissions:", round(mt_co2e, 2)), # Hover text
    highlightOptions = highlightOptions(
      weight = 3,
      color = "blue",
      bringToFront = TRUE
    )
  ) %>%
  addLegend(
    pal = pal,
    values = ~mt_co2e,
    title = "CTU Ag Emissions",
    opacity = 0.7
  )

# # Plot the emissions on the map - change this to leaflet later
# map_ctu_ag <- ggplot(data = ctu_ag_map) +
#   geom_sf(aes(fill = ctu_emissions)) +
#   scale_fill_viridis_c(option = "plasma") +  # You can change the color scale here
#   theme_minimal() +
#   labs(fill = "CTU Ag Emissions", title = "CTU Emissions Map")

fig_ctu_ag_emissions
```

## Summary
Methane production is largely tied to consumer food preferences, with enteric fermentation in cattle raised for meat and dairy production being the central driver of agricultural methane production. Nitrous oxide emissions also derive from livestock via manure emissions, both in managed systems and in manure that is applied to or leaches into soils. Crop production also accounts for a large percentage of agricultural nitrous oxide emissions via synthetic fertilizer production and the biochemical processes that delivers nitrogen to soils. Efficient agricultural techniques, particularly in crop management, can reduce the amount of nitrogen entering soil and therefore the amount that is volatilized into gaseous nitrous oxide.


```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
