## {Data Source}
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
hookaddcap()
```


Agricultural emissions are derived from two primary sources. First is the USDA agricultural census which provides county level livestock head counts, crop production, and fertilizer sales. Second is the EPA State Inventory Tool (SIT) that provides state specific values for emission factors as well as activity data such as state-wide fertilizer application (different from sales). These are both the highest rank of data.

The USDA census is conducted once every five years (years ending in '2 and '7). For interstitial years we used linear interpolation to fill in county level values.

The SIT provides guidance on how to translate livestock counts, crop production, and fertilizer use into emission factors. A brief description of the major emitters in this sector is listed here, but please refer to the SIT documentation for more detail information.

Enteric fermentation

Enteric fermentation is methane emitted from livestock during digestion.
Emissions (MMTCO2E) = Animal Population × Emission Factor (kg CH4/head) × 28 (GWP) ÷ 1,000,000,000 (kg/MMTCO2E)

Manure management

Manure in lagoons and holding facilities emits methane and nitrous oxide.

Manure runoff

Manure from managed and unmanaged systems runoff into soil and cause additional nitrous oxide emissions.


Cropland - Plant residue

Croplands emit nitrous oxide when via legumes (i.e. soybeans, alfalfa, beans) that fix atmospheric N2 in the soil, some of which is converted to nitrous oxide via soil biochemical processes. Additionally, non-legume crops have residues that are left on soils and breakdown into the soil, emitting further nitrous oxide.

Fertilizer application

Synthetic and organic fertilizers breakdown to provide N to crops, but some fraction of this is converted to nitrous oxide that is emitted to the atmosphere.

Fertilizer runoff

Some proportion of applied fertilized is leached into the soil and runs off into streams and adjacent soils, where it further is chemically converted to nitrous oxide.

Introduction text 
Data source description, type

- Quality rank (See @tbl-quality-rank)
- How, when, and why was the data collected?
- If this is a modeled dataset, what is the sample?
- What is the raw unit of measurement? 
- How was this data accessed? Include any relevant links/citations, code, or downloads.
- What data cleaning or wrangling was completed? How did you test these processes and outputs?
- What is the geographic and temporal scope? Did you complete any aggregation?
- What version is the data? Were there other versions available? If so, why did you choose this version?
- What assumptions are made when we use this dataset?
- Which subject matter expert (SME) reviewed this data?
- Describe testing used to verify data

Be sure to add a citation of this dataset to the Zotero shared library.

### Data characteristics

- Were there any missing data? How did you handle missing data?
- Plots, tables, and description of data distribution
- Variance, Z-Score, quantiles
- Facet views by categorical variables

```{r fig-scatter}
#| fig-cap: "A scatter plot"
#| out-width: "95%"
#| include: false
shinipsum::random_ggplotly(type = "point") %>%
  councilR::plotly_layout(
    main_title = "Scatter"
  )
```


```{r fig-boxplot}
#| fig-cap: "A boxplot"
#| out-width: "95%"
#| include: false
shinipsum::random_ggplotly(type = "boxplot") %>%
  councilR::plotly_layout(
    main_title = "Boxplot"
  )
```


```{r fig-density}
#| fig-cap: "A density plot"
#| out-width: "95%"
#| include: false
shinipsum::random_ggplotly(type = "density") %>%
  councilR::plotly_layout(
    main_title = "Density"
  )
```


### Limitations

### Comparison with similar datasets

### Data dictionary

Table with detailed description of columns and definitions for each data table.

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
