# Electricity {#sec-electricity}
```{r include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_leaflet_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

# ── Geography filter ──────────────────────────────────────────────────────────
seven_county <- c("Anoka", "Carver", "Dakota", "Hennepin", "Ramsey", "Scott", "Washington")

# ── Target year (single point of update) ─────────────────────────────────────
target_year <- 2022

# ── County-level electricity by sector ────────────────────────────────────────
county_elec <- readRDS(file.path(
  here::here(), "_energy/data/county_elec_emissions_by_sector.RDS"
)) %>%
  filter(county_name %in% seven_county)

# ── CTU-level electricity emissions ──────────────────────────────────────────
ctu_elec <- readRDS(file.path(
  here::here(), "_meta/data/ctu_emissions.RDS"
)) %>%
  filter(
    sector_alt == "Electricity"
  )

# ── CTU geometry and population ──────────────────────────────────────────────
cprg_ctu <- readRDS(file.path(here::here(), "_meta/data/cprg_ctu.RDS")) %>%
  filter(county_name %in% seven_county) %>%
  mutate(geometry = st_make_valid(geometry))

ctu_population <- readRDS(file.path(here::here(), "_meta/data/ctu_population.RDS"))

# ── Summary statistics ───────────────────────────────────────────────────────
elec_2022 <- county_elec %>%
  filter(emissions_year == target_year) %>%
  pull(value_emissions) %>%
  sum(na.rm = TRUE)

elec_2005 <- county_elec %>%
  filter(emissions_year == 2005) %>%
  pull(value_emissions) %>%
  sum(na.rm = TRUE)

res_elec_2022 <- county_elec %>%
  filter(emissions_year == target_year, sector == "Residential") %>%
  pull(value_emissions) %>%
  sum(na.rm = TRUE)

res_share <- res_elec_2022 / elec_2022

hookaddcap()
```

## Introduction

Electricity emissions are driven by the fuel mix used for power generation rather than by on-site combustion. The Twin Cities metro falls within the Midwest Reliability Organization West (MROW) subregion of EPA's Emissions & Generation Resource Integrated Database (eGRID), which provides the emissions factors used to convert electricity deliveries into CO~2~e. The MROW grid mix is shown in @tbl-elec-mrow-proportions.

```{r tbl-elec-mrow-proportions}
#| tbl-cap: "Grid Mix for MROW subregion of 2022 eGRID"
#| out-width: "95%"

eGRID_mix <- tibble(
  "Energy Source" = c("Coal", "Oil", "Gas", "Other Fossil", "Nuclear", "Hydro", "Biomass", "Wind", "Solar", "Geothermal", "Other"),
  Percentage = c(38.0, 0.1, 8.3, 0.1, 8.1, 3.9, 0.7, 39.8, 1.0, 0.0, 0.1) / 100
)

(eGRID_mix) %>%
  arrange(-Percentage) %>%
  gt() %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::fmt_percent(columns = "Percentage", decimals = 1) %>%
  gt::as_raw_html()
```

Emissions from electricity consumption have declined substantially since 2005, largely as a result of coal-to-wind transitions in the MROW grid. In 2023, Minnesota passed a bill mandating a carbon-free electricity standard by 2040, requiring 80% carbon-free retail sales by 2030 and 100% by 2040 for all electric utilities. This section presents emissions attributed to electricity delivered to residential, commercial, and industrial end users across the seven-county metro. Natural gas consumed at power plants for electricity generation is treated separately in the [Building Energy](#sec-building-energy) section's sector decomposition to avoid double counting.


## Results

### `r target_year` county and subsector breakdown

Electricity accounted for an estimated `r scales::comma(round(elec_2022), accuracy = 1)` metric tons CO~2~e across the seven-county metro in `r target_year`. Residential electricity accounts for approximately `r round(100 * res_share)`% of total electricity emissions.

```{r fig-elec-county-sector-emissions}
#| fig-cap: "County electricity emissions by sector"
#| out-width: "95%"
#| fig.height: 4
#| fig.width: 10

county_elec_yr <- county_elec %>%
  filter(emissions_year == target_year) %>%
  mutate(sector = factor(sector, levels = c(
    "Residential", "Commercial", "Industrial"
  )))

elec_sector_colors <- c(
  "Residential" = colors$councilBlue,
  "Commercial" = cprg_colors$cprg_green,
  "Industrial" = cprg_colors$cprg_yellow
)

fig_elec_county_sector <- plot_ly(
  type = "bar",
  source = "fig-elec-county-sector-emissions",
  data = county_elec_yr,
  y = ~ reorder(county_name, value_emissions),
  x = ~value_emissions,
  color = ~sector,
  colors = elec_sector_colors,
  hovertemplate = ~ paste0(
    county_name, " County<br>",
    sector, "<br>",
    round(value_emissions / 1000, digits = 1), " thousand metric tons CO<sub>2</sub>e<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = paste(target_year, "county electricity emissions by sector"),
    x_title = "Metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    barmode = "stack",
    legend = list(traceorder = "reversed"),
    hoverlabel = list(font = list(color = "black"))
  )

fig_elec_county_sector
```


### Baseline emissions

```{r fig-elec-baseline-emissions}
#| fig-cap: "Electricity emissions over time"
#| out-width: "95%"
#| fig.height: 4
#| fig.width: 8

elec_temporal <- county_elec %>%
  filter(emissions_year <= target_year) %>%
  group_by(emissions_year, sector) %>%
  summarize(value_emissions = sum(value_emissions, na.rm = TRUE), .groups = "drop") %>%
  mutate(sector = factor(sector, levels = c(
    "Residential", "Commercial", "Industrial"
  )))

fig_elec_baseline <- plot_ly(
  source = "fig-elec-baseline-emissions",
  data = elec_temporal,
  x = ~emissions_year,
  y = ~value_emissions,
  color = ~sector,
  colors = elec_sector_colors,
  type = "scatter",
  mode = "lines",
  stackgroup = "one",
  hovertemplate = ~ paste0(
    "Year: %{x}<br>",
    "Sector: %{text}<br>",
    "Emissions: %{y:,.0f} metric tons CO<sub>2</sub>e<br>",
    "<extra></extra>"
  ),
  text = ~sector
) %>%
  plotly_layout(
    main_title = "Electricity Emissions Over Time",
    x_title = "Year",
    y_title = "Metric tons CO<sub>2</sub>e",
    subtitle = "Seven-county metro area"
  ) %>%
  layout(
    hovermode = "x unified",
    hoverlabel = list(font = list(color = "black"))
  )

fig_elec_baseline
```


### CTU per-capita electricity emissions

Per-capita electricity emissions vary across the region, reflecting differences in housing density, building stock, commercial and industrial activity, and the extent to which a community's electricity demand is dominated by residential versus non-residential load.

```{r fig-ctu-elec-per-cap-map}
#| fig-cap: "Electricity emissions per capita (metric tons CO2e), by CTU"
#| out-width: "95%"
#| out-height: "500px"

ctu_elec_yr <- ctu_elec %>%
  filter(emissions_year == target_year) %>%
  group_by(geog_name, ctu_class) %>%
  summarize(value_emissions_elec = sum(value_emissions, na.rm = TRUE), .groups = "drop")

ctu_map_data <- cprg_ctu %>%
  st_drop_geometry() %>%
  distinct(ctu_name, ctu_class, gnis) %>%
  left_join(ctu_elec_yr, by = c("ctu_name" = "geog_name", "ctu_class")) %>%
  left_join(
    ctu_population %>%
      filter(inventory_year == target_year) %>%
      group_by(ctu_name, ctu_class) %>%
      summarize(ctu_population = sum(ctu_population), .groups = "drop"),
    by = c("ctu_name", "ctu_class")
  ) %>%
  mutate(
    emissions_per_cap = if_else(
      ctu_population > 0 & value_emissions_elec > 0,
      value_emissions_elec / ctu_population,
      NA_real_
    )
  ) %>%
  right_join(cprg_ctu %>% select(gnis, geometry), by = "gnis") %>%
  st_as_sf() %>%
  filter(!ctu_name == "Fort Snelling")

epc_range <- range(ctu_map_data$emissions_per_cap, na.rm = TRUE)

emissions_cap_pal <- colorNumeric(
  palette  = "Blues",
  domain   = sqrt(epc_range),
  na.color = "transparent"
)

legend_vals <- pretty(epc_range, n = 5)
legend_vals <- legend_vals[legend_vals >= 0 & legend_vals <= epc_range[2]]

fig_ctu_elec_per_cap_map <- leaflet(
  data    = ctu_map_data %>% filter(!is.na(emissions_per_cap)),
  options = leafletOptions(zoomSnap = 0.5)
) %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%
  addPolygons(
    fillColor = ~ emissions_cap_pal(sqrt(emissions_per_cap)),
    fillOpacity = 0.8,
    color = "white",
    weight = 0.5,
    smoothFactor = 0.5,
    label = ~ paste0(
      ctu_name, " (", ctu_class, ")<br>",
      "CO2e/capita: ", round(emissions_per_cap, 1), "<br>",
      "Population: ", scales::comma(ctu_population, accuracy = 1), "<br>",
      "Total CO2e: ", scales::comma(round(value_emissions_elec), accuracy = 1)
    ) %>% lapply(htmltools::HTML),
    labelOptions = labelOptions(
      style     = list("font-size" = "12px"),
      direction = "auto"
    )
  ) %>%
  addLegend(
    position = "bottomright",
    title    = "CO<sub>2</sub>e<br>per capita",
    colors   = emissions_cap_pal(sqrt(legend_vals)),
    labels   = round(legend_vals, 1),
    opacity  = 0.8
  )

fig_ctu_elec_per_cap_map
```


## Summary

Electricity is a major source of greenhouse gas emissions in the seven-county metro, but one that is declining as the MROW grid transitions from coal to wind and solar. The 2040 carbon-free electricity standard will accelerate this trend. Sector decomposition shows that residential electricity accounts for the largest share of emissions, with the non-residential remainder split into commercial and industrial using NREL SLOPE modeled ratios. At the CTU level, per-capita emissions are uneven across the region, spiking in communities with large commercial or industrial electricity loads relative to population. As building electrification increases — shifting heating and transportation loads onto the grid — continued grid decarbonization will be critical for realizing net emissions reductions.


```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>%
  stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
