# Electricity

```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

county_emissions_summary <- county_emissions %>%
  group_by(sector, category, source, data_source, factor_source) %>%
  summarize(
    value_emissions = sum(value_emissions),
    .groups = "keep"
  )

mn_emissions <- read_rds(file.path(here::here(), "_energy/data/minnesota_county_ElecEmissions.RDS"))
wi_emissions <- read_rds(file.path(here::here(), "_energy/data/wisconsin_county_ElecEmissions.RDS")) %>%
  rename(county = county_name)

elecEmissions_all <- county_emissions %>%
  filter(category == "Electricity") %>%
  group_by(county_name, emissions_year) %>%
  summarise(
    value_emissions = sum(value_emissions),
    .groups = "rowwise"
  )




eiaQA_all <- bind_rows(
  read_rds(
    file.path(
      here::here(),
      "_energy/data/minnesota_QA_versusEIAStateProfile.RDS"
    )
  ),
  read_rds(
    file.path(
      here::here(),
      "_energy/data/wisconsin_QA_versusEIAStateProfile.RDS"
    )
  )
)

EIAvMETC_tons <- eiaQA_all %>%
  mutate(
    downscaled_EIA_total_CO2e_emissions_tons = downscaled_EIA_total_CO2e_emissions_lbs / 2000,
    OURS_total_CO2e_emissions_tons = OURS_total_CO2e_emissions_lbs / 2000
  ) %>%
  pivot_longer(
    cols = c("downscaled_EIA_total_CO2e_emissions_tons", "OURS_total_CO2e_emissions_tons"),
    names_to = "inventory_source",
    values_to = "value_emissions"
  ) %>%
  mutate(
    inventory_source = recode(inventory_source,
      downscaled_EIA_total_CO2e_emissions_tons = "EIA",
      OURS_total_CO2e_emissions_tons = "Metropolitan Council"
    )
  )

nrelQA <- read_rds(
  file.path(
    here::here(),
    "_energy/data-raw/nrel_slope/nrel_emissions.RDS"
  )
) %>%
  filter(source == "Electricity" && year == 2021) %>%
  group_by(county_name) %>%
  summarise(total_value_emissions = sum(co2e))

nrel_metc_electricity_comp <- nrelQA %>%
  left_join(
    elecEmissions_all %>% select(county_name,
      MetC = value_emissions
    ),
    by = "county_name"
  ) %>%
  pivot_longer(
    cols = c("total_value_emissions", "MetC"),
    names_to = "inventory_source",
    values_to = "value_emissions"
  ) %>%
  mutate(
    inventory_source = recode(inventory_source,
      total_value_emissions =
        "NREL SLOPE",
      MetC =
        "Metropolitan Council"
    )
  )

electricity_2005 <- elecEmissions_all %>%
  filter(emissions_year == 2005) %>%
  pull(value_emissions) %>%
  sum()

electricity_2021 <- elecEmissions_all %>%
  filter(emissions_year == 2021) %>%
  pull(value_emissions) %>%
  sum()

hookaddcap()
```

## Introduction

Emissions from the electricity generation sector have declined by `r scales::percent((electricity_2005 - electricity_2021) / electricity_2005, accuracy = 0.1)` in the region since 2005, largely as a result of transitions in the grid towards energy sources such as wind and solar [@mpcaClimateChangeTrends2023]. In 2023, Minnesota Governor Walz signed a bill mandating a statewide carbon-free electricity standard by 2040. The law "establishes a standard for utilities to supply Minnesota customers with electricity generated or procured from carbon-free resources, beginning at an amount equal to 80% of retail sales for public utility customers in Minnesota in 2030 and increasing every 5 years to reach 100% for all electric utilities by 2040. The bill also requires that, by 2035, an amount equal to at least 55% of an electric utility’s total retail electric sales to customers in Minnesota must be generated or procured from eligible energy technologies." Wisconsin has not adopted a similar carbon-free electricity standard, but a Wisconsin DNR report noted both the economic gains from such to the renewable energy economy in the state, as well as the opportunities for decarbonization [@holtWisconsinCleanEnergy2019].


```{r tbl-elec-mrow-proportions}
#| tbl-cap: "Grid Mix for MROW subregion of 2021 eGRID"
#| out-width: "95%"

eGRID_mix <- data.frame(
  "Energy Source" = c("Coal", "Oil", "Gas", "Other Fossil", "Nuclear", "Hydro", "Biomass", "Wind", "Solar", "Geothermal", "Other", "Unknown/Purchased Fuel"),
  Percentage = c("39.6%", "0.2%", "10.6%", "0.10%", "8.6%", "4.4%", "0.8%", "34.6%", "0.9%", "0.0%", "0.2%", "N/A")
)

(eGRID_mix) %>%
  gt() %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::as_raw_html()
```

### Results
### 2021 county and subsector breakdown

```{r fig-energy-electricity-sector-emissions}
#| fig-cap: "2021 electricity emissions by sector"
#| out-width: "95%"


fig_energy_electricity_sector_emissions <-
  plot_ly(
    type = "bar",
    source = "fig-energy-electricity-sector-emissions",
    data = county_emissions %>%
      filter(
        emissions_year == 2021,
        category == "Electricity"
      ),
    y = ~ reorder(county_name, value_emissions),
    x = ~value_emissions,
    color = ~source,
    colors = unlist(category_colors),
    hovertemplate = ~ paste0(
      county_name, " County", "<br>",
      source, ", ", category, "<br>",
      round(value_emissions / 1000000, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "2021 electricity emissions by sector",
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Energy sector"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    )
  )
fig_energy_electricity_sector_emissions
```

```{r fig-electricity-baseline-emissions}
#| fig-cap: "Baseline electricity emissions"
#| out-width: "95%"
#| fig.height: 4  # Set height in inches
#| fig.width: 8   # Set width in inches

fig_electricity_baseline <- plot_ly(
  source = "fig-electricity-baseline-emissions",
  data = county_emissions %>%
    filter(category == "Electricity") %>%
    group_by(emissions_year, source) %>%
    summarise(value_emissions = sum(value_emissions)),
  x = ~emissions_year,
  y = ~value_emissions,
  color = ~source, #
  colors = unlist(category_colors), # Directly map the colors to categories
  type = "scatter",
  mode = "lines", # No points or lines, just fill
  stackgroup = "one", # Fill to the next y value (stacking)
  hoverinfo = "x+y+name", # Ensure hover displays x, y, and category name
  hovertemplate = ~ paste0(
    "Year: %{x}<br>",
    "Category: %{text}<br>",
    "Emissions: %{y:.2f} metric tons CO<sub>2</sub>e<br>",
    "<extra></extra>"
  ),
  text = ~source # Pass category to hovertext
) %>%
  plotly_layout(
    main_title = "Electricity Emissions Over Time",
    x_title = "Year",
    y_title = "Million metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    hovermode = "x unified", # Combine all hovertext under one hover
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_electricity_baseline
```

```{r fig-ctu-ag-emissions}
#| fig-cap: "2021 city level agricultural emissions"
#| out-width: "95%"
#| out-height: "500px"

# load the county boundaries layer

### load emissions and downscaling var (proportion of county ag land)
ctu_elec_emissions <- readRDS(file.path(here::here(), "./_energy/data/_ctu_electricity_emissions.rds"))

electricity_emissions_2021 <- ctu_elec_emissions %>%
  filter(inventory_year == 2021) %>%
  group_by(ctu_name, ctu_class) %>%
  summarize(mt_co2e = sum(value_emissions, na.rm = TRUE))

### map CTU scale agricultural emissions

cprg_ctu <- readRDS(here("_meta", "data", "cprg_ctu.RDS")) %>%
  mutate(geometry = st_make_valid(geometry))


ctu_elec_map <- left_join(cprg_ctu,
  electricity_emissions_2021,
  by = c("ctu_name", "ctu_class")
) %>%
  filter(
    !county_name %in% c(
      "St. Croix",
      "Pierce",
      "Sherburne",
      "Chisago"
    ),
    !is.na(county_name)
  ) %>%
  select(ctu_name, geometry, mt_co2e) %>%
  replace(is.na(.), 0)


# Define color palette based on the plasma scale from viridis
pal <- colorNumeric(palette = "plasma", domain = ctu_elec_map$mt_co2e)

# Create the leaflet map
map_ctu_elec <- leaflet(ctu_elec_map) %>%
  addTiles() %>% # Add default OpenStreetMap tiles
  addPolygons(
    fillColor = ~ pal(mt_co2e),
    color = "black", # Border color
    weight = 1,
    fillOpacity = 0.9, # Transparency of the fill color
    popup = ~ paste(ctu_name, " electricity emissions:", round(mt_co2e, 2)), # Popup with details
    label = ~ paste(ctu_name, " emissions:", round(mt_co2e, 2)), # Hover text
    highlightOptions = highlightOptions(
      weight = 3,
      color = "blue",
      bringToFront = TRUE
    )
  ) %>%
  addLegend(
    pal = pal,
    values = ~mt_co2e,
    title = "CTU Electricity Emissions",
    opacity = 0.7
  )


map_ctu_elec
```


```

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```

{{< pagebreak >}}
