# Electricity

```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

county_emissions_summary <- county_emissions %>%
  group_by(sector, category, source, data_source, factor_source) %>%
  summarize(
    value_emissions = sum(value_emissions),
    .groups = "keep"
  )

mn_emissions <- read_rds(file.path(here::here(), "_energy/data/minnesota_county_ElecEmissions.RDS"))
wi_emissions <- read_rds(file.path(here::here(), "_energy/data/wisconsin_county_ElecEmissions.RDS")) %>%
  rename(county = county_name)

elecEmissions_all <- county_emissions %>%
  filter(category == "Electricity") %>%
  group_by(county_name, emissions_year) %>%
  summarise(
    value_emissions = sum(value_emissions),
    .groups = "rowwise"
  )




eiaQA_all <- bind_rows(
  read_rds(
    file.path(
      here::here(),
      "_energy/data/minnesota_QA_versusEIAStateProfile.RDS"
    )
  ),
  read_rds(
    file.path(
      here::here(),
      "_energy/data/wisconsin_QA_versusEIAStateProfile.RDS"
    )
  )
)

EIAvMETC_tons <- eiaQA_all %>%
  mutate(
    downscaled_EIA_total_CO2e_emissions_tons = downscaled_EIA_total_CO2e_emissions_lbs / 2000,
    OURS_total_CO2e_emissions_tons = OURS_total_CO2e_emissions_lbs / 2000
  ) %>%
  pivot_longer(
    cols = c("downscaled_EIA_total_CO2e_emissions_tons", "OURS_total_CO2e_emissions_tons"),
    names_to = "inventory_source",
    values_to = "value_emissions"
  ) %>%
  mutate(
    inventory_source = recode(inventory_source,
      downscaled_EIA_total_CO2e_emissions_tons = "EIA",
      OURS_total_CO2e_emissions_tons = "Metropolitan Council"
    )
  )

nrelQA <- read_rds(
  file.path(
    here::here(),
    "_energy/data-raw/nrel_slope/nrel_emissions.RDS"
  )
) %>%
  filter(source == "Electricity" && year == 2021) %>%
  group_by(county_name) %>%
  summarise(total_value_emissions = sum(co2e))

nrel_metc_electricity_comp <- nrelQA %>%
  left_join(
    elecEmissions_all %>% select(county_name,
      MetC = value_emissions
    ),
    by = "county_name"
  ) %>%
  pivot_longer(
    cols = c("total_value_emissions", "MetC"),
    names_to = "inventory_source",
    values_to = "value_emissions"
  ) %>%
  mutate(
    inventory_source = recode(inventory_source,
      total_value_emissions =
        "NREL SLOPE",
      MetC =
        "Metropolitan Council"
    )
  )

electricity_2005 = elecEmissions_all %>% 
  filter(emissions_year == 2005) %>% 
  pull(value_emissions) %>% 
  sum()

electricity_2021 = elecEmissions_all %>% 
  filter(emissions_year == 2021) %>% 
  pull(value_emissions) %>% 
  sum()

hookaddcap()
```

## Introduction

Emissions from the electricity generation sector have declined by `r scales::percent((electricity_2005 - electricity_2021) / electricity_2005, accuracy = 0.1)` in the region since 2005, largely as a result of transitions in the grid towards energy sources such as wind and solar [@mpcaClimateChangeTrends2023]. In 2023, Minnesota Governor Walz signed a bill mandating a statewide carbon-free electricity standard by 2040. The law "establishes a standard for utilities to supply Minnesota customers with electricity generated or procured from carbon-free resources, beginning at an amount equal to 80% of retail sales for public utility customers in Minnesota in 2030 and increasing every 5 years to reach 100% for all electric utilities by 2040. The bill also requires that, by 2035, an amount equal to at least 55% of an electric utility’s total retail electric sales to customers in Minnesota must be generated or procured from eligible energy technologies." Wisconsin has not adopted a similar carbon-free electricity standard, but a Wisconsin DNR report noted both the economic gains from such to the renewable energy economy in the state, as well as the opportunities for decarbonization [@holtWisconsinCleanEnergy2019].

## Methods

The general workflow for quantifying electricity emissions is to identify all of the electric utilities that operate within our study area, collect any reporting they provide to the states of Minnesota and Wisconsin about the amount of energy delivered to all their customers (with reference to federal reporting sources where state-level reporting gaps exist), and apply EPA-provided emissions factors to the reported activity/energy deliveries to calculate estimated emissions. Methodologies for allocating utility activity reports to counties varies across MN and WI and are further described in the following section. Most inputs we use in the construction of our electricity emissions data set are of the highest quality rank (@tbl-quality-rank), as they are either direct government-created data (e.g., emissions factors) or data reported to state/federal authorities (e.g., regulatory filings). However, for two Minnesota electric utilities – Elk River Municipal Utilities [@elkrivermunicipalutilities2021AnnualFinancial2022] and New Prague Utilities Commission [@newpragueutilitiescommissionAgendaPacket242022] – where regulatory filing data could not be sourced to quantify electricity deliveries, we referred to financial reporting documents published by the utilities.

Total regional emissions (Emissions~r~) represents the sum of all recorded energy deliveries by utility *i* within county *j*, where *i* refers to each of the electric utilities operating across our region, and *j* refers to the eleven counties included in this inventory. Our regional total therefore represents an aggregation of electricity deliveries for all distinct utility-county records.

$$Emissions_r = \Sigma (mWhDelivered_iCounty_j \times {Emissions Factor}) $$

Our inventory takes a "demand-side" approach to emissions quantification and seeks to aggregate all reported delivery of energy to ALL customers served by utilities (meaning all customer types, inclusive of residential, commercial, industrial, and government accounts). This means that energy loss and use experienced by utilities in the process of energy generation and transmission, and delivery and resale to utilities operating outside of our study area, are *not* directly reflected in the numbers attributed to counties. The U.S. Energy Information Administration (EIA) estimates that annual electricity transmission and distribution (T&D) losses averaged about 5% of the electricity transmitted and distributed in the United States in 2018 through 2022. [@u.s.energyinformationadministrationHowMuchElectricity2023]

While our primary data collection does not include a breakout of electricity deliveries by sector, we do leverage year 2021 NREL SLOPE forecasts of electricity consumption (built from a base year of 2016 observed data) by sector (residential, commercial, industrial) at the county level to calculate modeled proportions of consumption by sector, which we then apply to our aggregate numbers to calculate estimated emissions *by sector* [@maSectorSpecificMethodologiesSubnatonal2019, @nrelSLOPEStateLocal2017].

### Identifying utilities in scope

To identify the electric utilities that operate within our 11-county study area, we referred to maps and geospatial datasets capturing utility service areas in Minnesota and Wisconsin. To identify Wisconsin electric utilities, we downloaded the Electric Service Territory map maintained by the Wisconsin Public Service Commission [@wisconsinpublicservicecommissionElectricServiceTerritories2024]. To identify Minnesota electric utilities, we downloaded the Electric Service Territory map maintained by the Minnesota Public Utilities Commission and the Minnesota IT Geospatial Information Office[@minnesotaitgeospatialinformationofficeElectricUtilityService2023].

### Collecting and aggregating activity data from utilities

#### County-level activity data

After identifying which utilities operate within our study area within each state, we collect reporting submitted by these utilities to the relevant state and federal authorities and use a variety of approaches, depending on data availability, to allocate utility activity/energy deliveries to specific counties. For the state of Minnesota, we collected reports provided by all in-scope utilities for years 2014-2023. For year 2021, additional manual data collection efforts were completed to guarantee a complete data set, since reports for all in-scope utilities were unfortunately not available for all years. As needed, data for other county-years could be finalized in a similar fashion.

##### Minnesota

All electric utilities authorized to do business in Minnesota are required to file an annual data report pursuant to MN Rules Chapter 7610. The Minnesota Public Utilities Commission makes these reports searchable through an [eFiling Site](https://www.edockets.state.mn.us/EFiling/edockets/searchDocuments.do?method=eDocketsResult&docketYear=22&docketNumber=19), and downloadable as Excel workbooks [@mndocCHAPTER7610ENERGY2005]. For each utility identified in [distinct_electricity_util_type_MN.RDS](data/distinct_electricity_util_type_MN.RDS) (a data product of [minnesota_electricUtilities.R](data-raw/minnesota_electricUtilities.R), a script that looks for intersections between electric utility service areas and our Minnesota counties), we downloaded the relevant 2021 annual reports from this site (see note about Great River Energy in the previous section for caveats), except for North Branch Municipal Water and Light, which did not submit a 2021 report (we used their 2022 report as a substitution). Elk River Municipal Utilities and New Prague Utilities Commission did not file reports for 2021 or 2022, and so we used financial reports to identify their total electricity delivered; both utilities operated within only one county in our study area, which meant no estimation/allocation was necessary.

We wrote code to extract the county-level data reported in the report section titled "ITS DELIVERIES TO ULTIMATE CONSUMERS BY COUNTY FOR THE LAST CALENDAR YEAR" on the relevant annual data report Excel workbooks compiled into a folder directory, and created a table with three columns: `county`, `utility`, and `mWh_delivered` (megawatt-hours). By compiling this data for all utilities found to operate within our study area, aggregating all electricity deliveries at the county level becomes possible.

##### Wisconsin

All municipal and investor-owned utilities authorized to do business in Wisconsin are required to file an annual report with financial and operational information pursuant to Wis. Stat. § 196.07. The Public Services Commission of Wisconsin makes these reports searchable through an [E-Services Portal](https://apps.psc.wi.gov/ARS/annualReports/default.aspx), and downloadable as either PDFs or Excel workbooks, with options to export only specific portions of the reports as spreadsheets [@publicservicecommwisconsinAnnualReportsView2022]. For each utility identified in [distinct_electricity_util_type_WI.RDS](data/distinct_electricity_util_type_WI.RDS) (a data product of [wisconsin_electricUtilities.R](_energy/data-raw/wisconsin_electricUtilities.R), a script that looks for intersections between electric utility service areas and our Wisconsin counties), we downloaded the relevant 2021 annual reports from this site.

A similar process was followed for the four Wisconsin cooperative utilities for which we referenced federal regulatory filings data to populate our dataset of electricity deliveries.

Because of the small amount of data, and the different data structures of the reports for municipally-, cooperatively-, and investor-owned utilities in Wisconsin, we hard-coded observed information into data frames rather than extracting data through a web scraper or document analyzer (see [processed_wi_electricUtil_activityData.R](data-raw/processed_wi_electricUtil_activityData.R)).

#### City-level activity data (seven-county metro in MN only)

No regulatory requirements in Minnesota or Wisconsin directly require utilities make city-level reports on electrical or natural gas usage data activity publicly available in a manner similar MN Rules Chapter 7610's requirement for public county-level reporting. In alignment with *climate planning statute*, the Met Council is leading an effort to collect this information on behalf of cities and townships in the seven-county region as a constituent part of the greenhouse gas inventories being developed by the Council to support communities in their own local climate planning as required by state statute and Imagine 2050 minimum requirements. This data collection effort entails 1) direct processing of real, city-sector level data provided by utilities in our region, as well as 2) a series of modeling efforts geared towards addressing missingness in our city/township dataset, as well as disaggregation of coarser numbers provided by utilities (e.g., combined data for commercial/industrial sectors) into residential, commercial, and industrial numbers.

##### Processing utility-provided data

Xcel, the largest electric utility in our region, publishes Community Energy Reports [@xcelenergy] that document energy deliveries to specific communities in their service area. We gathered this data for years 2015-2022 and compiled it into the final .RDS file [xcel_activityData_NREL_2015_2022_process.RDS](data/xcel_activityData_NREL_2015_2022_process.RDS).

[minnesota_xcelCommunityReports_electricity.R](_energy/data-raw/minnesota_xcelCommunityReports_electricity.R) includes all data processing steps necessary to process the nearly 1000 files (one for each city-year combination) made available on the Community Energy Reports site. The function `get_files` identifies and extracts relevant info (e.g., city name) from all data files, while `process_file` reads detailed electricity consumption data from sections titled “Standard Community Report” in each Excel workbook. Key metrics, such as kilowatt-hours delivered (`kWh_delivered`) and utility-reported carbon emissions (`util_reported_co2e`), are then transformed into aggregated datasets (`Xcel_activityData_2015_2023`) at the utility-year-county level.

Ambiguities in sector classifications, particularly the "Business" category, are addressed through proportional disaggregation using NREL-modeled emissions proportions. This process incorporates additional datasets, such as `nrel_slope_city_emission_proportions.RDS`, to allocate energy and emissions between commercial and industrial sectors. Geographic integration (i.e., disaggregating CTUs with footprint in multiple counties) is achieved using `cprg_ctu` and `ctu_population`, enabling population-based allocation of activity/utility-reported emissions to constituent COCTU units.

Final outputs of [minnesota_xcelCommunityReports_electricity.R](_energy/data-raw/minnesota_xcelCommunityReports_electricity.R) include clean, joined datasets (`xcel_activityData_NREL_2015_2022_process.RDS`) and visualizations that compare utility-reported sector proportions to modeled values. *More to say here before modeling missingness section?*

#### Modeling missingness

Content to come

#### Emissions factors

To transform electricity deliveries (recorded in mWh) into emissions in metric tons CO~2~e, we referenced Emissions & Generation Resource Integrated Database (eGRID) summary data for the MROW subregion (Midwest Reliability Organization - West) [@usepaEGRIDEmissionsGeneration2021].

This dataset provides estimates in lbs/mWh for CO~2~, CH~4~, N~2~O, and CO~2~e emissions based on MROW's sub-regional electricity generation grid mix; converting this factor to metric tons per mWh and multiplying the county-level mWh estimates yields an estimate for CO~2~e. To generate this sub-regional estimate, eGRID first calculates estimated emissions at the plant level, and assigns plants to regions. By using an emissions factor that is localized, our inventory accounts for the specific grid mix in our study area (see the grid mix linked to the eGRID MROW emissions factor used in this inventory below). Per the eGRID technical guide, "the subregions were defined to limit the import and export of electricity in order to establish an aggregated area where the determined emission rates most accurately matched the generation and emissions from the plants within that subregion."

```{r tbl-elec-mrow-proportions}
#| tbl-cap: "Grid Mix for MROW subregion of 2021 eGRID"
#| out-width: "95%"

eGRID_mix <- data.frame(
  "Energy Source" = c("Coal", "Oil", "Gas", "Other Fossil", "Nuclear", "Hydro", "Biomass", "Wind", "Solar", "Geothermal", "Other", "Unknown/Purchased Fuel"),
  Percentage = c("39.6%", "0.2%", "10.6%", "0.10%", "8.6%", "4.4%", "0.8%", "34.6%", "0.9%", "0.0%", "0.2%", "N/A")
)

(eGRID_mix) %>%
  gt() %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::as_raw_html()
```

### Results



```{r fig-energy-electricity-sector-emissions}
#| fig-cap: "2021 electricity emissions by sector"
#| out-width: "95%"


fig_energy_electricity_sector_emissions <-
  plot_ly(
    type = "bar",
    source = "fig-energy-electricity-sector-emissions",
    data = county_emissions %>%
      filter(emissions_year == 2021,
        category == "Electricity"),
    y = ~ reorder(county_name, value_emissions),
    x = ~value_emissions,
    color = ~source,
    colors = unlist(category_colors),
    hovertemplate = ~ paste0(
      county_name, " County", "<br>",
      source, ", ", category, "<br>",
      round(value_emissions / 1000000, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "2021 electricity emissions by sector",
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Energy sector"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    )
  )
fig_energy_electricity_sector_emissions
```

```{r fig-electricity-baseline-emissions}
#| fig-cap: "Baseline electricity emissions"
#| out-width: "95%"
#| fig.height: 4  # Set height in inches
#| fig.width: 8   # Set width in inches

fig_electricity_baseline <- plot_ly(
  source = "fig-electricity-baseline-emissions",
  data = county_emissions %>%
    filter(category == "Electricity") %>%
    group_by(emissions_year, source) %>%
    summarise(value_emissions = sum(value_emissions)),
  x = ~emissions_year,
  y = ~value_emissions,
  color = ~source, #
  colors = unlist(category_colors), # Directly map the colors to categories
  type = "scatter",
  mode = "lines", # No points or lines, just fill
  stackgroup = "one", # Fill to the next y value (stacking)
  hoverinfo = "x+y+name", # Ensure hover displays x, y, and category name
  hovertemplate = ~ paste0(
    "Year: %{x}<br>",
    "Category: %{text}<br>",
    "Emissions: %{y:.2f} metric tons CO<sub>2</sub>e<br>",
    "<extra></extra>"
  ),
  text = ~source # Pass category to hovertext
) %>%
  plotly_layout(
    main_title = "Electricity Emissions Over Time",
    x_title = "Year",
    y_title = "Million metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    hovermode = "x unified", # Combine all hovertext under one hover
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_electricity_baseline
```

#### Correlation with related data

We would expect counties with higher population to have higher emissions.

```{r fig-electricity-emissions-population}
#| fig-cap: "County population and electricity emissions"
#| out-width: "95%"
cprg_population <- readRDS(file.path(here::here(), "_meta/data/cprg_population.RDS"))

emissions_population <- left_join(elecEmissions_all, cprg_population,
  by = c("county_name")
) %>% 
  filter(emissions_year == 2021)

fig_electricity_emissions_population <- plot_ly(
  data = emissions_population,
  source = "fig-electricity-emissions-population",
  x = ~population,
  y = ~value_emissions,
  type = "scatter",
  mode = "markers",
  hovertemplate = ~ paste0(
    county_name, " County", "<br>",
    "Population: ", scales::comma(population), "<br>",
    round(value_emissions / 1000000, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  ),
  marker = list(
    size = 18,
    color = colors$councilBlue,
    line = list(
      color = "lightgray",
      width = 2
    )
  )
) %>%
  plotly_layout(
    main_title = "County population and emissions",
    subtitle = "2021 ACS 5-year population estimates",
    x_title = "Total population",
    y_title = "Metric tons CO<sub>2</sub>e"
  )

fig_electricity_emissions_population
```

#### Comparison with federal inventories

```{r region-total-electricity}
region_total_electricity <- county_emissions %>%
  ungroup() %>%
  filter(source == "Electricity",
         emissions_year == 2021) %>%
  summarize(value_emissions = sum(value_emissions)) %>%
  magrittr::extract2("value_emissions")

eia_total_electricity <- eiaQA_all$downscaled_EIA_total_CO2e_emissions_lbs %>%
  sum() %>%
  units::as_units("pound") %>%
  units::set_units("ton") %>%
  as.numeric()
```

<!--
CURRENTLY NOT WORKING
##### Energy Information Administration (EIA)

We compared our county-level emissions figures to emissions estimates derived from the EIA's State Electricity Profiles, which were generated by down-scaling EIA's Minnesota and Wisconsin state-level numbers for total electricity retail sales. We achieved this by applying the proportions of each state's population residing in each of our study area counties to generate county-level emissions estimates. Our inventory quantified `r scales::comma(region_total_electricity, accuracy = 1)` metric tons of electricity in our study area relative to `r scales::comma(eia_total_electricity, accuracy = 1)` metric tons based on the aforementioned downscaled statewide retail numbers.


```{r fig-eia-qa-natural-gas-plot}
#| fig-cap: "Metropolitan Council emissions inventory vs. downscaled EIA State Electricity Profiles"
#| out-width: "95%"

# Create the plotly graph
# fig_eia_qa_electricity_plot <- plot_ly(
#   data = EIAvMETC_tons,
#   x = ~ reorder(county_name, -value_emissions),
#   y = ~value_emissions,
#   source = "fig-eia-qa-natural-gas-plot",
#   color = ~inventory_source,
#   type = "bar",
#   colors = c(
#     "EIA" = cprg_colors$cprg_da_yellow,
#     "Metropolitan Council" = colors$councilBlue
#   ),
#   hovertemplate = ~ paste0(
#     county_name, "<br>",
#     inventory_source, "<br>",
#     round(value_emissions / 1000000, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
#     "<extra></extra>"
#   )
# ) %>%
#   plotly_layout(
#     main_title = "Electricity Emissions Inventory QA<br><sub>Comparison between Metropolitan Council and EIA State Electricity Profiles</sub>",
#     y_title = "Metric tons CO<sub>2</sub>e",
#     x_title = "County",
#     legend_title = "Inventory source"
#   ) %>%
#   layout(
#     barmode = "group",
#     hovermode = "closest",
#     legend = list(
#       traceorder = "normal"
#     )
#   )
# 
# # Show the plot
# fig_eia_qa_electricity_plot
```
-->

##### NREL SLOPE

```{r}
nrel_total_electrcity <- nrel_metc_electricity_comp %>%
  filter(inventory_source == "NREL SLOPE") %>%
  summarize(value_emissions = sum(value_emissions)) %>%
  extract2("value_emissions")
```

The NREL SLOPE (State and Local Planning for Energy) Platform provides yearly forecasted emissions tied to the user of electricity up to 2050 based on 2016 reported data at the county level. In comparing these figures to our own inventory, we observed that, where we estimated `r scales::comma(region_total_electricity, accuracy = 1)` metric tons of emissions linked to electricity deliveries in our study area in the year 2021, NREL SLOPE forecasted `r scales::comma(nrel_total_electrcity, accuracy = 1)`metrics tons in our study area.

```{r fig-nrel-qa-electricity}
#| fig-cap: "Metropolitan Council emissions inventory v. NREL SLOPE modeled emissions"
#| out-width: "95%"

# Create the plotly graph
fig_nrel_qa_electricity <- plot_ly(
  data = nrel_metc_electricity_comp %>% 
    filter(emissions_year == 2021),
  x = ~ reorder(county_name, -value_emissions),
  y = ~value_emissions,
  source = "fig-nrel-qa-electricity",
  color = ~inventory_source,
  type = "bar",
  colors = c(
    "NREL SLOPE" = cprg_colors$cprg_green,
    "Metropolitan Council" = colors$councilBlue
  ),
  hovertemplate = ~ paste0(
    county_name, "<br>",
    inventory_source, "<br>",
    round(value_emissions / 1000000, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = "Electricity Emissions Inventory QA<br><sub>Comparison between Metropolitan Council and NREL SLOPE</sub>",
    y_title = "Metric tons CO<sub>2</sub>e",
    x_title = "County",
    legend_title = "Inventory source",
  ) %>%
  layout(
    barmode = "group",
    hovermode = "closest",
    legend = list(
      traceorder = "normal"
    )
  )

# Show the plot
fig_nrel_qa_electricity
```


{{< pagebreak >}}
