# Building Energy {#sec-building-energy}
```{r include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_leaflet_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

# ── Geography filter ──────────────────────────────────────────────────────────
seven_county <- c("Anoka", "Carver", "Dakota", "Hennepin", "Ramsey", "Scott", "Washington")

# ── County-level natural gas by sector ────────────────────────────────────────
county_ng <- readRDS(file.path(
  here::here(), "_energy/data/county_natgas_emissions_by_sector.RDS"
)) %>%
  filter(county_name %in% seven_county)

county_ng_sector <- county_ng %>%
  filter(sector != "Powerplant")

county_ng_pp <- county_ng %>%
  filter(sector == "Powerplant")

# ── County-level propane and fuel oil emissions ───────────────────────────────
county_liquid <- readRDS(file.path(
  here::here(), "_energy/data/county_propane_fueloil_activity_emissions.RDS"
)) %>%
  filter(county_name %in% seven_county)

# ── CTU-level natural gas emissions ───────────────────────────────────────────
ctu_ng <- readRDS(file.path(
  here::here(), "_energy/data/_ctu_natgas_emissions.RDS"
)) %>%
  filter(county_name %in% seven_county)

# ── CTU-level liquid fuel emissions ───────────────────────────────────────────
ctu_liquid <- readRDS(file.path(
  here::here(), "_energy/data/_ctu_liquid_emissions.RDS"
)) %>%
  filter(county_name %in% seven_county)

# ── CTU geometry and population ───────────────────────────────────────────────
cprg_ctu <- readRDS(file.path(here::here(), "_meta/data/cprg_ctu.RDS")) %>%
  filter(county_name %in% seven_county) %>%
  mutate(geometry = st_make_valid(geometry))

ctu_population <- readRDS(file.path(here::here(), "_meta/data/ctu_population.RDS"))

# ── Conversion constants ──────────────────────────────────────────────────────
mcf_to_mmbtu <- 1.020

# ── Summary statistics ────────────────────────────────────────────────────────
target_year <- 2022

# Total building energy emissions in target year
# Natural gas sectors
ng_2022 <- county_ng_sector %>%
  filter(emissions_year == target_year) %>%
  pull(value_emissions) %>%
  sum(na.rm = TRUE)

ng_2005 <- county_ng_sector %>%
  filter(emissions_year == 2005) %>%
  pull(value_emissions) %>%
  sum(na.rm = TRUE)

# Liquid fuel total
liquid_2022 <- county_liquid %>%
  filter(emissions_year == target_year) %>%
  pull(value_emissions) %>%
  sum(na.rm = TRUE)

liquid_2005 <- county_liquid %>%
  filter(emissions_year == 2005) %>%
  pull(value_emissions) %>%
  sum(na.rm = TRUE)

building_energy_2022 <- ng_2022 + liquid_2022

building_energy_2005 <- ng_2005 + liquid_2005


# Residential share
res_ng_2022 <- county_ng_sector %>%
  filter(emissions_year == target_year, sector == "Residential") %>%
  pull(value_emissions) %>%
  sum(na.rm = TRUE)

res_share <- (res_ng_2022 + liquid_2022) / building_energy_2022

hookaddcap()
```

## Introduction

Building energy emissions encompass the combustion of natural gas, propane, and fuel oil for space heating, water heating, cooking, and other end uses in residential, commercial, and industrial buildings. Natural gas is the dominant fuel, delivered through the distribution networks of utilities operating under Minnesota Public Utilities Commission oversight. Propane and fuel oil serve communities beyond the reach of the gas distribution network, particularly in the region's rural areas.

This section presents emissions from all stationary fuel combustion within buildings. Electricity consumption, though a major component of building energy use, is treated separately in the [Electricity](#sec-electricity) section because its emissions derive from the generation mix rather than on-site combustion. Natural gas consumed for electric power generation at power plants is also treated separately; although power plant gas appears in utility delivery totals, those emissions are attributed to the electricity sector to avoid double counting.


## Results

### `r target_year` county and subsector breakdown

Building energy accounted for an estimated `r scales::comma(round(building_energy_2022), accuracy = 1)` metric tons CO~2~e across the seven-county metro in `r target_year`. Residential natural gas and liquid fuels account for approximately `r round(100 * res_share)`% of total building energy emissions.

```{r fig-be-county-sector-emissions}
#| fig-cap: "2022 county building energy emissions by sector"
#| out-width: "95%"
#| fig.height: 4
#| fig.width: 10

# Combine NG sectors and liquid fuels for the stacked bar
county_be_2022 <- bind_rows(
  county_ng_sector %>%
    filter(emissions_year == target_year) %>%
    transmute(county_name, sector, value_emissions),
  county_liquid %>%
    filter(emissions_year == target_year) %>%
    group_by(county_name) %>%
    summarize(value_emissions = sum(value_emissions, na.rm = TRUE), .groups = "drop") %>%
    mutate(sector = "Propane & fuel oil")
)

# Order sectors for stacking
county_be_2022 <- county_be_2022 %>%
  mutate(sector = factor(sector, levels = c(
    "Residential", "Business", "Industrial", "Power plant", "Propane & fuel oil"
  )))

be_sector_colors <- c(
  "Residential" = colors$councilBlue,
  "Business" = cprg_colors$cprg_green,
  "Industrial" = cprg_colors$cprg_yellow,
  "Power plant" = cprg_colors$cprg_da_yellow,
  "Propane & fuel oil" = cprg_colors$epa_blue
)

fig_be_county_sector <- plot_ly(
  type = "bar",
  source = "fig-be-county-sector-emissions",
  data = county_be_2022,
  y = ~ reorder(county_name, value_emissions),
  x = ~value_emissions,
  color = ~sector,
  colors = be_sector_colors,
  hovertemplate = ~ paste0(
    county_name, " County<br>",
    sector, "<br>",
    round(value_emissions / 1000, digits = 1), " thousand metric tons CO<sub>2</sub>e<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = paste(target_year, "county building energy emissions by sector"),
    x_title = "Metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    barmode = "stack",
    legend = list(traceorder = "reversed"),
    hoverlabel = list(font = list(color = "black"))
  )

fig_be_county_sector
```


### Baseline emissions

```{r fig-be-baseline-emissions}
#| fig-cap: "Building energy emissions over time"
#| out-width: "95%"
#| fig.height: 4
#| fig.width: 8

# Temporal stacked area — natural gas sectors + liquid fuels
ng_temporal <- county_ng_sector %>%
  filter(emissions_year <= target_year) %>%
  group_by(emissions_year, sector) %>%
  summarize(value_emissions = sum(value_emissions, na.rm = TRUE), .groups = "drop")

liquid_temporal <- county_liquid %>%
  filter(emissions_year <= target_year) %>%
  group_by(emissions_year) %>%
  summarize(value_emissions = sum(value_emissions, na.rm = TRUE), .groups = "drop") %>%
  mutate(sector = "Propane & fuel oil")

be_temporal <- bind_rows(ng_temporal, liquid_temporal) %>%
  mutate(sector = factor(sector, levels = c(
    "Residential", "Business", "Industrial", "Propane & fuel oil"
  )))

fig_be_baseline <- plot_ly(
  source = "fig-be-baseline-emissions",
  data = be_temporal,
  x = ~emissions_year,
  y = ~value_emissions,
  color = ~sector,
  colors = be_sector_colors,
  type = "scatter",
  mode = "lines",
  stackgroup = "one",
  hovertemplate = ~ paste0(
    "Year: %{x}<br>",
    "Sector: %{text}<br>",
    "Emissions: %{y:,.0f} metric tons CO<sub>2</sub>e<br>",
    "<extra></extra>"
  ),
  text = ~sector
) %>%
  plotly_layout(
    main_title = "Building Energy Emissions Over Time",
    x_title = "Year",
    y_title = "Metric tons CO<sub>2</sub>e",
    subtitle = "Seven-county metro area"
  ) %>%
  layout(
    hovermode = "x unified",
    hoverlabel = list(font = list(color = "black"))
  )

fig_be_baseline
```


### CTU per-capita energy intensity

Total residential heating fuel emissions per person (CO2e per capita, combining natural gas, propane, and fuel oil) varies substantially across the region. Higher per-capita values typically reflect lower-density housing, older building stock, and greater reliance on propane or fuel oil. This metric helps identify communities where demand-side efficiency programs and electrification may have the greatest impact.

```{r fig-ctu-mmbtu-per-cap-map}
#| fig-cap: "Residential heating fuel per capita (mmBtu), by CTU"
#| out-width: "95%"
#| out-height: "500px"

# ── Aggregate CTU-level NG to mmBtu ──────────────────────────────────────────
ctu_ng_yr <- ctu_ng %>%
  filter(
    emissions_year == target_year,
    sector == "Residential"
  ) %>%
  group_by(ctu_name, ctu_class) %>%
  summarize(value_emissions_ng = sum(value_emissions, na.rm = TRUE), .groups = "drop")

# ── Propane + fuel oil ────────────────────────────────────────────────────────
ctu_liquid_yr <- ctu_liquid %>%
  filter(emissions_year == max(emissions_year[emissions_year <= target_year])) %>%
  group_by(ctu_name, ctu_class) %>%
  summarize(value_emissions_fo = sum(value_emissions, na.rm = TRUE), .groups = "drop")

# ── Join with geometry and population ─────────────────────────────────────────
ctu_map_data <- cprg_ctu %>%
  st_drop_geometry() %>%
  distinct(ctu_name, ctu_class, gnis) %>%
  left_join(ctu_ng_yr, by = c("ctu_name", "ctu_class")) %>%
  left_join(ctu_liquid_yr, by = c("ctu_name", "ctu_class")) %>%
  left_join(
    ctu_population %>%
      filter(inventory_year == target_year) %>%
      group_by(ctu_name, ctu_class) %>%
      summarize(ctu_population = sum(ctu_population), .groups = "drop"),
    by = c("ctu_name", "ctu_class")
  ) %>%
  mutate(
    total_emissions = value_emissions_ng + value_emissions_fo,
    emissions_per_cap = if_else(
      ctu_population > 0 & total_emissions > 0,
      total_emissions / ctu_population,
      NA_real_
    )
  ) %>%
  right_join(cprg_ctu %>% select(gnis, geometry), by = "gnis") %>%
  st_as_sf() %>%
  filter(!ctu_name == "Fort Snelling")

epc_range <- range(ctu_map_data$emissions_per_cap, na.rm = TRUE)

emissions_cap_pal <- colorNumeric(
  palette  = "Blues",
  domain   = sqrt(epc_range),
  na.color = "transparent"
)

# Legend breakpoints in original units
legend_vals <- pretty(epc_range, n = 5)
legend_vals <- legend_vals[legend_vals >= 0 & legend_vals <= epc_range[2]]

fig_ctu_emissions_per_cap_map <- leaflet(
  data    = ctu_map_data %>% filter(!is.na(emissions_per_cap)),
  options = leafletOptions(zoomSnap = 0.5)
) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    fillColor = ~ emissions_cap_pal(sqrt(emissions_per_cap)),
    fillOpacity = 0.8,
    color = "white",
    weight = 0.5,
    smoothFactor = 0.5,
    label = ~ paste0(
      ctu_name, " (", ctu_class, ")<br>",
      "Total CO2e/capita: ", round(emissions_per_cap, 1), "<br>",
      "Population: ", scales::comma(ctu_population, accuracy = 1), "<br>",
      "Total CO2e: ", scales::comma(round(total_emissions), accuracy = 1)
    ) %>% lapply(htmltools::HTML),
    labelOptions = labelOptions(
      style     = list("font-size" = "12px"),
      direction = "auto"
    )
  ) %>%
  addLegend(
    position = "bottomright",
    title    = "CO<sub>2</sub>e<br>per capita",
    colors   = emissions_cap_pal(sqrt(legend_vals)),
    labels   = round(legend_vals, 1),
    opacity  = 0.8
  )

fig_ctu_emissions_per_cap_map
```


## Summary

Natural gas is the dominant source of building energy emissions in the seven-county metro, with residential consumption accounting for the largest share. The commercial sector is estimated as the residual of county utility deliveries after subtracting residential, industrial, and power plant gas — an approach that likely includes some smaller industrial users. Propane and fuel oil are a small share of regional totals but account for large shares of heating energy in rural cities and townships beyond the gas distribution network. Per-capita energy intensity is highest in lower-density communities with older housing stock, suggesting that building retrofit and electrification programs targeted at these areas could yield outsized emissions reductions and reduction of energy cost burdens.


```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>%
  stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
