## Propane and kerosene
```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r,include = FALSE}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

fuel_use <- readRDS(file.path(here::here(), "_energy/data/fuel_use.RDS")) %>%
  rename(
    county_name = NAME,
    emissions_year = year,
    value_emissions = emissions_metric_tons_co2e
  )

county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

county_emissions_summary <- county_emissions %>%
  group_by(sector, category, source, data_source, factor_source) %>%
  summarize(
    value_emissions = sum(value_emissions),
    .groups = "keep"
  )
hookaddcap()
```

[U.S Energy Information Administration](www.eia.gov) provides per household estimates of residential energy use from residential surveys of propane and kerosene usage patterns and billing data at regional and state scales. These estimates are provided every 5 years, most recently in 2020 [@useia2020ResidentialEnergy2023]. [The American Community Survey](https://www.census.gov/programs-surveys/acs), curated by the US Census Bureau, provides the estimated number of households using each fuel type at more granular spatial and temporal scales, including county level estimates in 2021. Energy generation and GHG emissions from these fuels was estimated for each of the 11 counties in our 2021 inventory using these two data sources in combination. Both data sources are from federal governmental agencies, categorized as the highest rank of data quality (@tbl-quality-rank). 

EIA RECS data was downloaded from their [data portal](https://www.eia.gov/consumption/residential/data/2020/). EIA surveyed 32,882 households nationally for 2020 data and provides estimated mmBtu (millions of British thermal units) generation from propane per household using that fuel at the state level and at the Midwest regional level for kerosene. ACS data was accessed using the tidycensus package in R. ACS conducted 62,778 household interviews in Minnesota in 2021 and provides county estimates  of the number of households using a given fuel. Multiplying the estimated mmBtu generated per household in a given region by the county-level estimate of households using the fuel provides a county-level estimate of mmBtu generation from each of propane and kerosene and is then converting  to CO2 equivalency using the EPA's emission factors. This approach assumes energy generation per household per fuel is equal across each state (for propane) and the Midwest region (for kerosene).

```{r fig-energy-liquid-stationary-fuel-emissions}
#| fig-cap: "2021 annual liquid stationary fuel emissions"
#| out-width: "95%"

fig_energy_liquid_stationary_fuel_emissions <-
  plot_ly(
    type = "bar",
    source = "fig-energy-liquid-stationary-fuel-emissions",
    data = fuel_use,
    y = ~ reorder(county_name, value_emissions),
    x = ~value_emissions,
    color = ~fuel_type,
    colors = unlist(source_colors),
    hovertemplate = ~ paste0(
      county_name, " County", "<br>",
      fuel_type, "<br>",
      round(value_emissions / 1000, digits = 2), " thousand metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "2021 annual liquid stationary fuel emissions",
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Liquid fuel type"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    )
  )

fig_energy_liquid_stationary_fuel_emissions
```




```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```

{{< pagebreak >}}
