# Natural gas

```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_leaflet_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
hookaddcap()
```

## Utility Activity Data

### Minnesota

Under Minnesota Administrative Rules Chapter 7610 [@mndocCHAPTER7610ENERGY2005], utilities are required to file an annual data report that supports the identification of "emerging energy trends based on supply and demand, conservation and public health and safety factors, and to determine the level of statewide and service area needs." [@minnesotadepartmentofcommerceAnnualReportingForms2022] This includes a report of county-level energy deliveries (reported in thousand cubic feet, commonly written as mcf). Because the information is structured in this manner, natural gas emissions at the county-level can be estimated as a direct function of energy deliveries to counties reported by utilities, which isn't the case in Wisconsin (some modeling and estimation is required in WI).

### Wisconsin

Under Wis. Stat. ยง 196.07, investor-owned natural gas utilities operating within the state of Wisconsin must submit annual reports to the State which include an array of information related to utility finance and operations, including key figures leveraged in our data collection, such as total energy deliveries made (in units of *therms*) and total number of customer accounts within each county [@wisconsinstatelegChapter196Regulation2024]. Because all four "in-scope" natural gas utilities in Wisconsin are investor-owned, we did not have to refer to federal sources or use population figures to estimate energy deliveries. We estimated county-wide emissions from natural gas in 2021 by first calculating the proportion of each utility customer accounts that were linked to either Pierce or St. Croix counties, and allocating that proportion of the utility's total reported energy delivery to each county. This approach represents a divergence from our Minnesota process, which involves aggregating county-level numbers directly reported by utilities, and implicitly assumes that customer accounts across counties within the operations of a given utility have the same average per-account demand for energy, when in actuality this is likely impacted by land-use mix and relative magnitude/scale of residential and commercial/industrial utility accounts within a given county.

$$mcfDelivered_iCounty_j =  (mcfEnergyDeliveries_i \times {ProportionOfTotalUtilityCustomers_j}) $$

Note: This approach implicitly assumes that customer accounts across counties within the operations of a given utility have the same average per-account demand for energy, when this is influenced by land-use mix and relative magnitude/scale of residential and commercial/industrial utility accounts within a given county.

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```

{{< pagebreak >}}
