# Electricity

```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_leaflet_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
hookaddcap()
```

## Utility Activity Data

### Minnesota

Under Minnesota Administrative Rules Chapter 7610 [@mndocCHAPTER7610ENERGY2005], utilities are required to file an annual data report that supports the identification of "emerging energy trends based on supply and demand, conservation and public health and safety factors, and to determine the level of statewide and service area needs" [@minnesotadepartmentofcommerceAnnualReportingForms2022]. This includes a report of county-level energy deliveries (reported in megawatt-hours, commonly written as mWh). Because the information is structured in this manner, electricity emissions at the county-level can be estimated as a direct function of energy deliveries to counties reported by utilities.

One utility operating within our study area, Great River Energy (GRE), is a non-profit wholesale electric power cooperative which provides wholesale power and delivery services to 28 Minnesota-based electric cooperatives, which collectively own GRE and provide retail electric service. They are the primary supplier of energy (\>99%) to the cooperative utilities operating in our study area. Though most electricity suppliers having relationships with Minnesota electric utilities do not file these annual data reports, GRE does, given their unique hybrid wholesale/transmission structure and cooperative ownership model. As a result, while we keep only GRE's reporting in our data collection for county-level electricity deliveries, we exclude reporting from these subsidiary retail electric cooperatives, in order to avoid double counting electric deliveries. We use GRE's reported deliveries to our 9 Minnesota study area counties as a full substitution for the deliveries of these utilities, which may represent a marginal undercount given that marginal-to-negligible amounts of energy delivered by these retail cooperatives came from other other sources, per their annual reports.

### Wisconsin

Under Wis. Stat. ยง 196.07, investor- and municipally-owned electric utilities operating within the state of Wisconsin must submit annual reports to the State which include an array of information related to utility finance and operations, including key figures leveraged in our data collection, such as total energy deliveries made (in units of *kWh*) and total number of customer accounts within each county [@wisconsinstatelegChapter196Regulation2024].

Of the seven in-scope electric utilities, only three (the investor- and municipally-owned utilities) were required to make these reports to the State in 2021 (four of the in-scope electric utilities are cooperative utilities); state data was leveraged in this case. For the four utilities (all cooperative utilities) not making these reports, we relied upon the detailed data files provided by the EIA, recording the responses of utilities to the Annual Electric Power Energy Report (Form EIA-861), which records retail sales by utilities and power marketers [@eia861]. Two utilities (Dunn Energy Cooperative and Polk-Burnett Electric Cooperative) filled the long version of the form (filled by larger entities), and two (St. Croix Electric Cooperative and Pierce-Pepin Electric Cooperative Services) filled the short form (filled by smaller entities). For our purposes, both the long and short form provided suitable activity information (total energy delivered, total customers) to allocate energy deliveries to counties in concert with Census population data, in the process outlined below.

Because Wisconsin utilities do not report energy deliveries at the county level, it was necessary to estimate energy deliveries by a given utility *i* within a particular county *j.* For those three utilities who reported county-level customer counts and total customer counts to the state, we estimated mWh delivered by utility *i* in county *j* by multiplying their total statewide energy deliveries (as reported to the relevant state authorities) by the proportion of their customers residing in each of our two study area counties.

$$mWhDelivered_iCounty_j =  (mWhEnergyDeliveries_i \times {ProportionOfTotalUtilityCustomers_j}) $$

Note: This approach implicitly assumes that customer accounts across counties within the operations of a given utility have the same average per-account demand for energy, when this is influenced by land-use mix and relative magnitude/scale of residential and commercial/industrial utility accounts within a given county.

To calculate the estimated energy delivered by utility *i* in county *j* for the four cooperatively-owned utilities that did not report county-level customer counts (i.e., did not report to the State), we used population figures to estimate/allocate reported electricity deliveries to counties. We took the actual total energy delivered by utility *i* across Wisconsin (as reported to the relevant federal authorities) and multiplied this by the proportion of total population within each utility's entire service area residing within county *j* at the 2020 decennial Census.

$$mWhDelivered_iCounty_j =  (mWhDelivered_i \times {ProportionOfTotalUtilityPopulation_j}) $$

The factor *ProportionOfTotalUtilityPopulation*~j~ was calculated by spatially joining Census block centroids containing population data to 1) polygons representing the entirety of our in-scope utilities' service areas (including areas outside of St. Croix and Pierce counties) and 2) polygons representing only the portions of these utilities' service areas within a) St. Croix county and b) Pierce County. These utility-county service areas are calculated separately, to facilitate an informed allocation of statewide energy use to each county in turn.

### Data dictionaries

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```

{{< pagebreak >}}
