# Methods and data sources  

```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
cprg_population <- readRDS(file.path(here::here(), "_meta/data/cprg_population.RDS"))

hookaddcap()
```

# Solid waste

## Methods

{{< include methods_solid_waste.qmd >}}

## Data sources

{{< include data_solid_waste.qmd >}}

<!-- data_wi_emissions.qmd is linked within the above file. -->

# Wastewater

## Methods

## Data sources

{{< include data_wastewater.qmd >}}

## Data Validation

### Correlation with related data

We would expect counties with a higher population to have higher solid waste emissions.

```{r fig-emissions-population-solidwaste}
#| fig-cap: "County population and solid waste emissions"
#| out-width: "95%"
cprg_population <- readRDS(file.path(here::here(), "_meta/data/cprg_population.RDS"))

solidwaste_totals <- county_emissions %>%
  filter(category == "Solid waste") %>%
  group_by(geog_name) %>%
  summarise(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)
  )

solidwaste_population <- left_join(solidwaste_totals, cprg_population,
  by = c("geog_name" = "NAME")
)

fig_emissions_population_solidwaste <- plot_ly(
  data = solidwaste_population,
  source = "fig-emissions-population-solidwaste",
  x = ~population,
  y = ~emissions_metric_tons_co2e,
  type = "scatter",
  mode = "markers",
  hovertemplate = ~ paste0(
    geog_name, " County", "<br>",
    "Population: ", scales::comma(population), "<br>",
    "Emissions: ", round(emissions_metric_tons_co2e * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  ),
  # opacity = 0.7,
  marker = list(
    color = colors$councilBlue,
    size = 18,
    line = list(
      color = "lightgray",
      width = 2
    )
  )
) %>%
  plotly_layout(
    main_title = "County population and solid waste emissions",
    subtitle = "2021 ACS 5-year population estimates. Strong positive correlation",
    x_title = "Total population",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "",
    legend = list(orientation = "h")
  )

fig_emissions_population_solidwaste
```

## Comparison with other inventories {#sec-waste-inventory-comparison}

#### US Greenhouse Gas Emissions Inventory

The United States EPA conducts a comprehensive yearly estimate of greenhouse gas emissions from multiple sectors and gases. It also publishes statewide totals consistent with the national inventory. These emissions totals are consistent with international standards for greenhouse gas accounting, although they may differ from inventories completed at the state level for various reasons.

US Inventory data for the waste sector in both Minnesota and Wisconsin was downloaded from the [Greenhouse Gas Inventory Data Explorer](https://cfpub.epa.gov/ghgdata/inventoryexplorer/) and processed in R script: [epa_inventory_data.R](/data-raw/epa_inventory_data.R), where it was apportioned from state to county level by population.

```{r fig-federal-inventory-compare}
#| fig-cap:  "Solid waste emissions comparison: Met Council and US GHG Inventory"
#| out-width: "95%"

waste_federal <- solidwaste_totals %>%
  mutate(data_source = "Met Council") %>%
  bind_rows(federal_totals)

fig_federal_inventory_compare <- plot_ly(
  data = waste_federal,
  source = "fig-federal-inventory-compare",
  x = ~geog_name,
  y = ~emissions_metric_tons_co2e,
  color = ~data_source,
  type = "bar",
  colors = c(
    "Met Council" = colors$councilBlue,
    "US GHG Inventory" = cprg_colors$cprg_da_yellow
  ),
  hovertemplate = ~ paste0(
    geog_name, " County", "<br>",
    data_source, "<br>",
    round(emissions_metric_tons_co2e * 1e-3, digits = 0),
    " thousand metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  councilR::plotly_layout(
    main_title = "Solid waste emissions comparison",
    subtitle = "Met Council estimates compared to US GHG Inventory",
    x_title = "County",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Emissions Source"
  ) %>%
  plotly::layout(barmode = "group")

fig_federal_inventory_compare
```


```{r metc-solid-epa-comp}
metc_solidwaste_inventory <- county_emissions %>%
  filter(sector == "Waste") %>%
  mutate(
    data_source = "Met Council",
    year = as.character(year)
  ) %>%
  select(names(epa_waste_inventory))

metc_epa_comp_solid <- epa_waste_inventory %>%
  filter(source != "Total") %>%
  bind_rows(metc_solidwaste_inventory) %>%
  mutate(
    source_unified = case_when(
      # source == "Anaerobic digestion" ~ "Waste to energy",
      TRUE ~ source
    )
  ) %>%
  select(-source)

wide_metc_epa_comp_solid <- metc_epa_comp_solid %>%
  pivot_wider(
    names_from = data_source,
    values_from = emissions_metric_tons_co2e
  ) %>%
  filter(
    !is.na(`Met Council`),
    !is.na(`US GHG Inventory`)
  ) %>%
  mutate(emiss_diff = `US GHG Inventory` - `Met Council`)
```

```{r plot_waste_comparison}
plot_waste_comparison <- function(waste_source,
                                  plotly_source) {
  plot_ly(
    source = plotly_source,
    data = metc_epa_comp_solid %>%
      filter(
        source_unified == waste_source
      ),
    y = ~geog_name,
    x = ~emissions_metric_tons_co2e,
    color = ~data_source,
    colors = c(
      "Met Council" = colors$councilBlue,
      "US GHG Inventory" = cprg_colors$cprg_da_yellow
    ),
    hovertemplate = ~ paste0(
      geog_name, " County", "<br>",
      data_source, "<br>",
      source_unified, "<br>",
      round(emissions_metric_tons_co2e * 1e-3, digits = 0), " thousand metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    ),
    marker = list(size = 18),
    opacity = 0.75,
    type = "scatter",
    mode = "markers"
  ) %>%
    add_segments(
      name = "Difference line",
      data = wide_metc_epa_comp_solid %>%
        filter(source_unified == waste_source),
      y = ~geog_name,
      yend = ~geog_name,
      x = ~`US GHG Inventory`,
      xend = ~`Met Council`,
      showlegend = FALSE,
      inherit = FALSE,
      line = list(
        color = "darkgray",
        size = 2
      )
    ) %>%
    councilR::plotly_layout(
      main_title = paste0(waste_source, " emissions comparison"),
      subtitle = "Met Council vs. US Greenhouse Gas Inventory estimates",
      y_title = "County",
      x_title = "Metric tons CO<sub>2</sub>e",
      legend_title = "Data source"
    ) %>%
    layout(yaxis = list(categoryorder = "total ascending"))
}
```

Wastewater comparisons are very consistent, with the US GHG Inventory consistently estimating more emissions than the EPA derived Met Council estimate. 

```{r fig-metc-epa-ghg-wastewater}
#| fig-cap: "Wastewater emissions comparison by county, EPA-derived Met Council and US GHG Inventory. US inventory consistently higher than Council. "
#| out-width: "95%"

fig_metc_epa_ghg_wastewater <- plot_waste_comparison("Wastewater",
  plotly_source = "fig-metc-epa-ghg-wastewater"
)
fig_metc_epa_ghg_wastewater
```

Solid waste emissions differences vary by county; while Ramsey, Hennepin and Washington county estimates from the US GHG Inventory are higher than the Council estimates, the opposite is true for Dakota and Anoka counties. This is likely related to the fact that Ramsey, Hennepin and Washington counties process a significant portion of their waste in waste-to-energy facilities, reducing the amount of emissions accounted for under the landfill category.

```{r fig-metc-epa-ghg-landfill}
#| fig-cap: "Landfill emissions comparison by county, Met Council and US GHG Inventory. Results vary by county."
#| out-width: "95%"

fig_metc_epa_ghg_landfill <- plot_waste_comparison("Landfill",
  plotly_source = "fig-metc-epa-ghg-landfill"
)
fig_metc_epa_ghg_landfill
```

```{r fig-solid-waste-versions-compare}
#| fig-cap:  "Solid waste emissions comparison: Met Council and US GHG Inventory"
#| out-width: "95%"

solidwaste_totals_prev <- waste_emissions_old %>%
  group_by(geoid) %>%
  summarise(
    value_emissions = sum(value_emissions)
  )

solidwaste_totals_new <- solid_waste %>%
  filter(inventory_year == 2021) %>%
  group_by(geoid) %>%
  summarise(
    value_emissions = sum(value_emissions)
  ) %>%
  mutate(data_source = "Met Council V2")

waste_federal <- solidwaste_totals_prev %>%
  mutate(data_source = "Met Council V1") %>%
  bind_rows(solidwaste_totals_new) %>%
  bind_rows(federal_totals)

fig_sw_versions_compare <- plot_ly(
  data = waste_federal,
  source = "fig-solid-waste-versions-compare",
  x = ~geoid,
  y = ~value_emissions,
  color = ~data_source,
  type = "bar",
  colors = c(
    "Met Council V1" = colors$councilBlue,
    "Met Council V2" = cprg_colors$cprg_green,
    "US GHG Inventory" = cprg_colors$cprg_da_yellow
  ),
  hovertemplate = ~ paste0(
    geoid, " County", "<br>",
    data_source, "<br>",
    round(value_emissions * 1e-3, digits = 0),
    " thousand metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  councilR::plotly_layout(
    main_title = "Solid waste emissions comparison",
    subtitle = "Met Council estimates compared to US GHG Inventory",
    x_title = "County",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Emissions Source"
  ) %>%
  plotly::layout(barmode = "group")

fig_sw_versions_compare
```


{{< pagebreak >}}

