# Waste and wastewater {#sec-waste}
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
source(file.path(here::here(), "R/global_warming_potential.R"))

cprg_population <- readRDS(file.path(here::here(), "_meta/data/cprg_population.RDS"))
cprg_county <- readRDS(file.path(
  here::here(),
  "_meta/data/cprg_county.RDS"
))

cprg_ctu <- readRDS(file.path(here::here(), "_meta/data/cprg_ctu.RDS"))
county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))
solid_waste_gas <- readRDS(file.path(here::here(), "_waste/data/solid_waste_MN_by_gas.RDS")) %>%
  left_join(cprg_county %>% st_drop_geometry() %>%
    select(geoid, county_name))

wastewater_gas <- readRDS(file.path(here::here(), "_waste/data/_county_wastewater_emissions.RDS"))

# this file is currently out of date. please update before adding full sector graphs


emissions_2021 <- county_emissions %>%
  filter(emissions_year == 2021) %>%
  pull(value_emissions) %>%
  sum()
waste_emissions_2021 <- county_emissions %>%
  filter(emissions_year == 2021, sector == "Waste") %>%
  pull(value_emissions) %>%
  sum()
waste_emissions_2005 <- county_emissions %>%
  filter(emissions_year == 2005, sector == "Waste") %>%
  pull(value_emissions) %>%
  sum()

county_gas_type <- bind_rows(
  wastewater_gas %>%
    group_by(inventory_year, county_name, units_emissions) %>%
    summarize(value_emissions = sum(value_emissions)),
  solid_waste_gas %>%
    group_by(inventory_year, county_name, units_emissions) %>%
    summarize(value_emissions = sum(value_emissions))
) %>%
  ungroup() %>%
  mutate(co2e = case_when(
    units_emissions == "Metric tons CH4" ~ value_emissions * gwp$ch4,
    units_emissions == "Metric tons N2O" ~ value_emissions * gwp$n2o,
    units_emissions == "Metric tons CO2" ~ value_emissions * gwp$co2
  ))

co2_2021 <- county_gas_type %>%
  filter(
    inventory_year == 2021,
    units_emissions == "Metric tons CO2"
  ) %>%
  pull(co2e) %>%
  sum()

methane_2021 <- county_gas_type %>%
  filter(
    inventory_year == 2021,
    units_emissions == "Metric tons CH4"
  ) %>%
  pull(co2e) %>%
  sum()

n2o_2021 <- county_gas_type %>%
  filter(
    inventory_year == 2021,
    units_emissions == "Metric tons N2O"
  ) %>%
  pull(co2e) %>%
  sum()


hookaddcap()
```

## Introduction

In 2021, waste activities over the entire United States generated emissions of 169.2 MMTCO~2~e, or 2.7 percent of total U.S. greenhouse gas emissions [@usepaInventoryGreenhouseGas2023]. In the 11-county Twin Cities MSA region, municipal waste produced `r 100*round(waste_emissions_2021)` MMTCO~2~e, or `r 100*round(waste_emissions_2021/emissions_2021, digits = 3)`% of the regional total. 

Municipal waste can broadly be divided into biogenic wastewater emissions and solid waste emissions. Our inventory accounts for industrial waste emission point sources in the industrial sector and wastewater processing emissions such as electricity and transportation in their respective sectors. Note that the Metropolitan Council Climate Action Work Plan accounts for holistic operational inventory emissions from wastewater processing.

## Results

```{r fig-waste-county-emissions}
#| fig-cap: "2021 county agricultural emissions"
#| out-width: "95%"
#| fig.height: 4  # Set height in inches
#| fig.width: 10   # Set width in inches

fig_waste_county_emissions <- plot_ly(
  type = "bar",
  source = "fig-waste-county-emissions",
  data = county_emissions %>%
    filter(
      sector == "Waste",
      emissions_year == 2021
    ),
  y = ~ reorder(county_name, value_emissions),
  x = ~value_emissions,
  color = ~category,
  colors = unlist(category_colors),
  hovertemplate = ~ paste0(
    county_name, "<br>",
    category, "<br>",
    source, "<br>",
    round(value_emissions / 1000000, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = "2021 county waste emissions",
    x_title = "Million metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_waste_county_emissions
```

### Baseline emissions

Regional waste emissions have increased by `r 100*round(1 - (waste_emissions_2021/waste_emissions_2005))`% since 2005, likely tied to population growth despite successful measures to reduce emissions per capita.

```{r fig-waste-baseline-emissions}
#| fig-cap: "Baseline agricultural emissions"
#| out-width: "95%"
#| fig.height: 4  # Set height in inches
#| fig.width: 8   # Set width in inches

fig_waste_baseline <- plot_ly(
  source = "fig-waste-baseline-emissions",
  data = county_emissions %>%
    filter(sector == "Waste") %>%
    group_by(emissions_year, category) %>%
    summarise(value_emissions = sum(value_emissions)),
  x = ~emissions_year,
  y = ~value_emissions,
  color = ~category, #
  colors = unlist(category_colors), # Directly map the colors to categories
  type = "scatter",
  mode = "lines", # No points or lines, just fill
  stackgroup = "one", # Fill to the next y value (stacking)
  hoverinfo = "x+y+name", # Ensure hover displays x, y, and category name
  hovertemplate = ~ paste0(
    "Year: %{x}<br>",
    "Category: %{text}<br>",
    "Emissions: %{y:.2f} metric tons CO<sub>2</sub>e<br>",
    "<extra></extra>"
  ),
  text = ~category # Pass category to hovertext
) %>%
  plotly_layout(
    main_title = "County Waste Emissions Over Time",
    x_title = "Year",
    y_title = "Million metric tons CO<sub>2</sub>e",
    subtitle = ""
  ) %>%
  layout(
    hovermode = "x unified", # Combine all hovertext under one hover
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_waste_baseline
```

### Emissions by gas type

The majority of waste emissions by weight are CO<sub>2</sub>, but when accounting for the increased warming potential of methane, CO<sub>2</sub>e, methane contributes the most to global warming. 
CO<sub>2</sub>: `r 100 * round(co2_2021/waste_emissions_2021, digits = 3)`% 
CH<sub>4</sub>: `r 100 * round(methane_2021/waste_emissions_2021, digits = 3)`% 
N<sub>2</sub>O: `r 100 * round(n2o_2021/waste_emissions_2021, digits = 3)`%



```{r fig-waste-county-emissions-by-gas-type}
#| fig-cap: "2021 county waste emissions by gas type"
#| out-width: "95%"

fig_waste_gas_type <- plot_ly(
  type = "bar",
  source = "fig-waste-county-emissions-by-gas-type",
  data = county_gas_type %>%
    filter(inventory_year == 2021),
  y = ~ reorder(county_name, value_emissions),
  x = ~value_emissions,
  color = ~units_emissions,
  hovertemplate = ~ paste0(
    county_name, "<br>",
    units_emissions, "<br>",
    round(value_emissions, digits = 2), "  metric tons of gas", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = "2021 county waste emissions by gas type",
    x_title = "Metric tons of gas emitted",
    subtitle = ""
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_waste_gas_type
```
```{r fig-waste-county-emissions-by-gas-type-co2e}
#| fig-cap: "2021 county waste emissions by gas type - CO2e"
#| out-width: "95%"

fig_waste_gas_type_co2e <- plot_ly(
  type = "bar",
  source = "fig-waste-county-emissions-by-gas-type-co2e",
  data = county_gas_type %>%
    filter(inventory_year == 2021),
  y = ~ reorder(county_name, co2e),
  x = ~co2e,
  color = ~units_emissions,
  hovertemplate = ~ paste0(
    county_name, "<br>",
    units_emissions, "<br>",
    round(co2e, digits = 2), "  metric tons of gas", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = "2021 county waste emissions by gas type in CO2e",
    x_title = "Metric tons of CO2e",
    subtitle = ""
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    hoverlabel = list(
      font = list(
        color = "black" # Set hovertext color to black
      )
    )
  )

fig_waste_gas_type_co2e
```

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
