# Waste and wastewater 
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
cprg_population <- readRDS(file.path(here::here(), "_meta/data/cprg_population.RDS"))
epa_waste_inventory <- readRDS(file.path(here::here(), "_waste/data-raw/epa_solidwaste_inventory.RDS"))

federal_totals <- epa_waste_inventory %>%
  filter(source == "Total") %>%
  select(
    geog_name,
    source,
    emissions_metric_tons_co2e,
    data_source
  )

cprg_ctu <- readRDS(file.path(here::here(), "_meta/data/cprg_ctu.RDS"))
county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

county_emissions_summary <- county_emissions %>%
  group_by(sector, category, source, data_source, factor_source) %>%
  summarize(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e),
    .groups = "keep"
  )

sector_percentages <- county_emissions_summary %>%
  group_by(sector) %>%
  summarize(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)) %>%
  arrange(-emissions_metric_tons_co2e) %>%
  janitor::adorn_percentages(denominator = "col")


waste_percentages <- county_emissions_summary %>%
  filter(sector == "Waste") %>% 
  group_by(category, source) %>%
  summarize(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e), .groups = "keep") %>%
  arrange(-emissions_metric_tons_co2e) %>%
  janitor::adorn_percentages(denominator = "col")
hookaddcap()
```

## Introduction

In 2021, waste activities generated emissions of 169.2 MMTCO~2~e, or 2.7 percent of total U.S. greenhouse gas emissions [@usepaInventoryGreenhouseGas2023]. Solid waste emissions primarily consist of CH~4~, emitted in large part by landfills and organics composting, but also include CO~2~ and N~2~O generation. In the metropolitan area, waste incineration accounts for a significant portion of solid waste management and therefore emissions. This inventory also includes emissions generated in the management of recyclables.


Wastewater emissions account for direct CH~4~ and N~2~O emissions from the treatment of municipal wastewater, and not additional indirect anthropogenic emissions such as the electricity needed to operate processing plants.

## Methods {#sec-waste-methods}

The process for calculating solid waste emissions slightly differs between Minnesota and Wisconsin counties, due to a difference in data availability.

For Minnesota counties, waste generation totals are allocated by sector from Minnesota Pollution Control Agency (MPCA) SCORE data [@mpcaMPCASCOREGuidebook2022]. Totals for each sector are multiplied by the appropriate EPA-provided emissions factor to provide an estimate of emissions from that source and county. Learn more about MPCA SCORE in @sec-mpca-score. 

For Wisconsin counties, state-level emissions data as calculated by the Wisconsin DNR is allocated to the relevant counties by population [@wisconsindnrGreenhouseGas2021].

The emissions for each county $c$ are proportional to its fractional share of Wisconsin's population.

$$Emissions_c = Emissions_{total} \times \frac{Pop_c}{Pop_{total}}$$

## Results

Emissions from waste is the smallest of the three sectors in the Twin Cities MSA. Waste generated `r round(sum(county_emissions$emissions_metric_tons_co2e)/1000000 * sector_percentages[[2]][3], digits = 2)` MMtCO~2~e of emissions in the Twin Cities MSA in 2021. Solid waste, including landfills, recycling, and organics, generates the largest share of emissions in the waste sector, with wastewater treatment comprising a smaller share of waste emissions. 

```{r fig-waste-wastewater-county-emissions}
#| fig-cap: "Solid waste and wastewater county emissions"
#| out-width: "95%"
fig_waste_wastewater_county_emissions <-
  plot_county_sector_emissions(county_emissions,
                               .sector = "Waste",
                               .plotly_source = "fig-waste-wastewater-county-emissions"
  )

fig_waste_wastewater_county_emissions$x$layoutAttrs[[1]]$title$text <- "Solid waste and wastewater emissions"

fig_waste_wastewater_county_emissions
```


```{r fig-waste-county-emissions-by-category}
#| fig-cap: "Solid waste and wastewater county emissions by category"
#| out-width: "95%"

fig_waste_county_emissions_by_category <- plot_county_emissions(
  county_emissions = county_emissions,
  .sector = "Waste",
  .plotly_source = "fig-waste-county-emissions-by-category"
)

fig_waste_county_emissions_by_category$x$layoutAttrs[[1]]$title$text <- "Solid waste and wastewater emissions"

fig_waste_county_emissions_by_category
```

### Wastewater

```{r wastewater-percentages, include = FALSE}
total_wastewater <- 
  county_emissions %>% 
  filter(category == "Wastewater") %>% 
  ungroup() %>% 
  summarise(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e), 
            .groups = "keep") %>% 
  extract2("emissions_metric_tons_co2e")
```

Wastewater generated `r round(total_wastewater/1000000, digits = 2)` MMtCO~2~e of emissions in the Twin Cities MSA in 2021.

```{r fig-wastewater-sector-emissions}
#| fig-cap: "2021 wastewater emissions"
#| out-width: "95%"

fig_wastewater_sector_emissions <-
  plot_ly(
    type = "bar",
    source = "fig-wastewater-sector-emissions",
    data = county_emissions %>%
      filter(source == "Wastewater"),
    y = ~ reorder(geog_name, emissions_metric_tons_co2e),
    x = ~emissions_metric_tons_co2e,
    color = ~category,
    colors = unlist(category_colors),
    hovertemplate = ~ paste0(
      geog_name, " County", "<br>",
      category, "<br>",
      round(emissions_metric_tons_co2e / 1000, digits = 0), " thousand metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "2021 wastewater emissions",
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = ""
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    yaxis = list(categoryorder = "total ascending")
  )
fig_wastewater_sector_emissions
```


### Solid waste

```{r solid-waste-percentages, include=FALSE}
solid_waste_percentages <- county_emissions %>% 
  filter(category == "Solid waste") %>% 
  group_by(category, source) %>% 
  summarize(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)) %>% 
  arrange(-emissions_metric_tons_co2e) %>%
  janitor::adorn_percentages(denominator = "col")
  

total_solid_waste <- 
  county_emissions %>% 
  filter(category == "Solid waste") %>% 
  ungroup() %>% 
  summarise(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)) %>% 
  extract2("emissions_metric_tons_co2e")
```

Solid waste generated `r round(total_solid_waste/1000000, digits = 2)` MMtCO~2~e of emissions in the Twin Cities MSA in 2021. Of that total, `r scales::percent(solid_waste_percentages$emissions_metric_tons_co2e[1], accuracy = 0.1)` of emissions came from `r str_to_lower(solid_waste_percentages$source[1])`, `r scales::percent(solid_waste_percentages$emissions_metric_tons_co2e[2], accuracy = 0.1)` from `r str_to_lower(solid_waste_percentages$source[2])` facilities, and the remaining `r scales::percent(sum(solid_waste_percentages$emissions_metric_tons_co2e[3:4]), accuracy = 0.1)` from `r str_to_lower(solid_waste_percentages$source[3])` and `r str_to_lower(solid_waste_percentages$source[4])`.  

```{r fig-solid-waste-emissions}
#| fig-cap: "2021 solid waste emissions"
#| out-width: "95%"
fig_solid_waste_emissions <-
  plot_ly(
    type = "bar",
    source = "fig-solid-waste-emissions",
    data = county_emissions %>%
      filter(category == "Solid waste"),
    y = ~ reorder(geog_name, emissions_metric_tons_co2e),
    x = ~emissions_metric_tons_co2e,
    color = ~category,
    colors = unlist(category_colors),
    marker = list(
      line = list(
        color = colors$councilBlue
      )
    ),
    hovertemplate = ~ paste0(
      geog_name, " County", "<br>",
      category, ", ", source, "<br>",
      round(emissions_metric_tons_co2e / 1000, digits = 0), " thousand metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "2021 solid waste emissions",
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Category"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    yaxis = list(categoryorder = "total ascending")
  )
fig_solid_waste_emissions
```

Greenhouse gas emissions from solid waste are dominated by the landfill sector. In Hennepin, Ramsey, and Washington counties, municipal centers where a significant portion of waste is incinerated, waste-to-energy or incineration makes up a large fraction of emissions as well. 

```{r fig-solid-waste-sector-emissions-by-category}
#| fig-cap: "2021 solid waste emissions by category"
#| out-width: "95%"

fig_solid_waste_sector_emissions_by_category <-
  plot_ly(
    type = "bar",
    source = "fig-solid-waste-sector-emissions-by-category",
    data = county_emissions %>%
      filter(category == "Solid waste"),
    y = ~ reorder(geog_name, emissions_metric_tons_co2e),
    x = ~emissions_metric_tons_co2e,
    color = ~source,
    colors = unlist(source_colors),
    hovertemplate = ~ paste0(
      geog_name, " County", "<br>",
      category, "<br>",
      round(emissions_metric_tons_co2e / 1000, digits = 0), " thousand metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "2021 solid waste emissions by category",
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Category"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    yaxis = list(categoryorder = "total ascending")
  )
fig_solid_waste_sector_emissions_by_category
```


## Correlation with related data

We would expect counties with a higher population to have high wastewater emissions. The resulting correlation is strong. 

```{r fig-emissions-population-wastewater}
#| fig-cap: "County population and wastewater emissions"
#| out-width: "95%"

wastewater_totals <- county_emissions %>%
  filter(source == "Wastewater") %>% 
  group_by(geog_name) %>%
  summarise(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)
  )

wastewater_population <- left_join(wastewater_totals, cprg_population,
                                   by = c("geog_name" = "NAME")
)

fig_emissions_population_wastewater <- plot_ly(
  data = wastewater_population,
  source = "fig-emissions-population-wastewater",
  x = ~population,
  y = ~emissions_metric_tons_co2e,
  type = "scatter",
  mode = "markers",
  hovertemplate = ~ paste0(
    geog_name, " County", "<br>",
    "Population: ", scales::comma(population), "<br>",
    "Emissions: ", round(emissions_metric_tons_co2e * 1e-3, digits = 0),
    " thousand metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  ),
  # opacity = 0.7,
  marker = list(
    color = colors$councilBlue,
    size = 18,
    line = list(
      color = "lightgray",
      width = 2
    )
  )
) %>%
  plotly_layout(
    main_title = "County population and wastewater emissions",
    subtitle = "2021 ACS 5-year population estimates. Strong positive correlation",
    x_title = "Total population",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "",
    legend = list(orientation = "h")
  )

fig_emissions_population_wastewater
```


We would expect counties with a higher population to have high solid waste emissions.

```{r fig-emissions-population-solidwaste}
#| fig-cap: "County population and solid waste emissions"
#| out-width: "95%"
cprg_population <- readRDS(file.path(here::here(), "_meta/data/cprg_population.RDS"))

solidwaste_totals <- county_emissions %>%
  filter(category == "Solid waste") %>% 
  group_by(geog_name) %>%
  summarise(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)
  )

solidwaste_population <- left_join(solidwaste_totals, cprg_population,
                                   by = c("geog_name" = "NAME")
)

fig_emissions_population_solidwaste <- plot_ly(
  data = solidwaste_population,
  source = "fig-emissions-population-solidwaste",
  x = ~population,
  y = ~emissions_metric_tons_co2e,
  type = "scatter",
  mode = "markers",
  hovertemplate = ~ paste0(
    geog_name, " County", "<br>",
    "Population: ", scales::comma(population), "<br>",
    "Emissions: ", round(emissions_metric_tons_co2e * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  ),
  # opacity = 0.7,
  marker = list(
    color = colors$councilBlue,
    size = 18,
    line = list(
      color = "lightgray",
      width = 2
    )
  )
) %>%
  plotly_layout(
    main_title = "County population and solid waste emissions",
    subtitle = "2021 ACS 5-year population estimates. Strong positive correlation",
    x_title = "Total population",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "",
    legend = list(orientation = "h")
  )

fig_emissions_population_solidwaste
```

## Comparison with other inventories {#sec-waste-inventory-comparison}

#### US Greenhouse Gas Emissions Inventory

The United States EPA conducts a comprehensive yearly estimate of greenhouse gas emissions from multiple sectors and gases. It also publishes statewide totals consistent with the national inventory. These emissions totals are consistent with international standards for greenhouse gas accounting, although they may differ from inventories completed at the state level for various reasons.

US Inventory data for the waste sector in both Minnesota and Wisconsin was downloaded from the [Greenhouse Gas Inventory Data Explorer](https://cfpub.epa.gov/ghgdata/inventoryexplorer/) and processed in R script: [epa_inventory_data.R](/data-raw/epa_inventory_data.R), where it was disaggregated from state to county level by population.

```{r fig-federal-inventory-compare}
#| fig-cap:  "Solid waste emissions comparison: Met Council and US GHG Inventory"
#| out-width: "95%"

waste_federal <- solidwaste_totals %>%
  mutate(data_source = "Met Council") %>%
  bind_rows(federal_totals)

fig_federal_inventory_compare <- plot_ly(
  data = waste_federal,
  source = opts_current$get()$label,
  x = ~geog_name,
  y = ~emissions_metric_tons_co2e,
  color = ~data_source,
  type = "bar",
  colors = c(
    "Met Council" = colors$councilBlue,
    "US GHG Inventory" = cprg_colors$cprg_da_yellow
  ),
  hovertemplate = ~ paste0(
    geog_name, " County", "<br>",
    data_source, "<br>",
    round(emissions_metric_tons_co2e * 1e-3, digits = 0),
    " thousand metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  councilR::plotly_layout(
    main_title = "Solid waste emissions comparison",
    subtitle = "Met Council estimates compared to US GHG Inventory",
    x_title = "County",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Emissions Source"
  ) %>%
  plotly::layout(barmode = "group")

fig_federal_inventory_compare
```


```{r metc-solid-epa-comp}
metc_solidwaste_inventory <- county_emissions %>%
  filter(sector == "Waste") %>%
  mutate(
    data_source = "Met Council",
    year = as.character(year)
  ) %>%
  select(names(epa_waste_inventory))

metc_epa_comp_solid <- epa_waste_inventory %>%
  filter(source != "Total") %>%
  bind_rows(metc_solidwaste_inventory) %>%
  mutate(
    source_unified = case_when(
      # source == "Anaerobic digestion" ~ "Waste to energy",
      TRUE ~ source
    )
  ) %>%
  select(-source)

wide_metc_epa_comp_solid <- metc_epa_comp_solid %>%
  pivot_wider(
    names_from = data_source,
    values_from = emissions_metric_tons_co2e
  ) %>%
  filter(
    !is.na(`Met Council`),
    !is.na(`US GHG Inventory`)
  ) %>% 
  mutate(emiss_diff = `US GHG Inventory` - `Met Council`)

```

```{r plot_waste_comparison}
plot_waste_comparison <- function(waste_source,
                                  plotly_source){
  plot_ly(
    source = plotly_source,
    data = metc_epa_comp_solid %>%
      filter(
        source_unified == waste_source
      ),
    y = ~geog_name,
    x = ~emissions_metric_tons_co2e,
    color = ~data_source,
    colors = c(
      "Met Council" = colors$councilBlue,
      "US GHG Inventory" = cprg_colors$cprg_da_yellow
    ),
    hovertemplate = ~ paste0(
      geog_name, " County", "<br>",
      data_source, "<br>",
      source_unified, "<br>",
      round(emissions_metric_tons_co2e * 1e-3, digits = 0), " thousand metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    ),
    marker = list(size = 18),
    opacity = 0.75,
    type = "scatter",
    mode = "markers"
  ) %>%
    add_segments(
      name = "Difference line",
      data = wide_metc_epa_comp_solid %>%
        filter(source_unified == waste_source),
      y = ~geog_name,
      yend = ~geog_name,
      x = ~`US GHG Inventory`,
      xend = ~`Met Council`,
      showlegend = FALSE,
      inherit = FALSE,
      line = list(
        color = "darkgray",
        size = 2
      )
    ) %>%
    councilR::plotly_layout(
      main_title = paste0(waste_source, " emissions comparison"),
      subtitle = "Met Council vs. US Greenhouse Gas Inventory estimates",
      y_title = "County",
      x_title = "Metric tons CO<sub>2</sub>e",
      legend_title = "Data source"
    ) %>%
    layout(yaxis = list(categoryorder = "total ascending"))
}
```

Wastewater comparisons are very consistent, with the US GHG Inventory consistently estimating more emissions than the Met Council estimate. 


```{r fig-metc-epa-ghg-wastewater}
#| fig-cap: "Wastewater emissions comparison by county, Met Council and US GHG Inventory. US inventory consistently higher than Council. "
#| out-width: "95%" 

fig_metc_epa_ghg_wastewater <- plot_waste_comparison("Wastewater",
                                                     plotly_source = "fig-metc-epa-ghg-wastewater")
fig_metc_epa_ghg_wastewater
```

Solid waste emissions differences vary by county; while Ramsey and Hennepin county estimates from the US GHG Inventory are higher than the Council estimates, the opposite is true for Dakota and Anoka counties. 

```{r fig-metc-epa-ghg-landfill}
#| fig-cap: "Landfill emissions comparison by county, Met Council and US GHG Inventory. Results vary by county."
#| out-width: "95%" 

fig_metc_epa_ghg_landfill <- plot_waste_comparison("Landfill",
                                                   plotly_source = "fig-metc-epa-ghg-landfill")
fig_metc_epa_ghg_landfill
```


```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
