
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
hookaddcap()
```

<!-- {{< include data_mpca_score.qmd >}} -->

### MPCA Waste Characterization Study

In 2013, the MPCA contracted with Burns & McDonnell Engineering Company, Inc. to conduct a waste characterization study of landfill waste, an update to an earlier study in the year 2000. For simplicity, we have chosen to set waste proportions equal to those 2013 values for all years in this inventory.

The study sampled waste makeup from six waste disposal facilities across the state and used the results to model statewide totals. More details can be found in [study documentation](https://www.pca.state.mn.us/sites/default/files/w-sw1-60.pdf).

This data is a high-quality state-level dataset.

For consistency, the values in this study have been compared to the IPCC default waste breakdown for North America. Relevant values were found to be consistent within 90% confidence intervals (as reported in the MPCA study), with the exception of food waste. IPCC values attribute 33.9% of landfill waste to food waste, while the 2013 report only finds that 17.8% of landfill waste is food waste. There are many possible explanations for this discrepancy, including differences in waste breakdown across the North American region and the possibility that the IPCC's numbers reflect additional food waste processed in organics facilities, which would not be included in the MPCA study.

Due to the fact that it is more recent and more specific to the Minnesota region, we have chosen the MPCA study as the source of truth in this case.

<!-- If necessary, display final breakdown of percentages aligned with DOC calculation categories here. -->

{{< include data_wi_emissions.qmd >}}

### EPA Methane Recovery Data {#sec-epa-methane}

The EPA generates methane flaring and landfill gas to energy data for each state as part of its State Inventory Tool for solid waste. This data is collected as part of the [Landfill Methane Outreach Program](https://www.epa.gov/lmop). 

However, since this data is collected on a national level, there are potential discrepancies with state-level and especially regional breakdowns of methane recovery. Due to these concerns, we have chosen to exclude this data source from our inventory and instead use the IPCC default of 0 methane recovery. This means that our inventory may overestimate emissions from solid waste landfills.


```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
