## Wastewater
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
hookaddcap()
```

We currently have three sources for wastewater emissions: the EPA's state inventory tool, the MPCA's emissions inventory, and the Met Council's Environmental Services emission estimates. Because only the EPA's estimate extends to Wisconsin and Minnesota we are reporting those data for consistency, but explore all three below where they have geographic overlap. The EPA estimate uses state population, biochemical oxygen demand, fraction of wastewater anaerobically digested, fraction of population on septic, estimated protein content, and biosolids used as fertilizer to estimate CH4 and N2O emissions. These inputs are modifiable and should be explored further in the CCAP. We down-scaled the state level emission estimates to county levels by population percentage.

- Data is the highest quality rank (federal agency)
- Data is modeled by the agencies based on multiple parameters, including population, aerobic vs anaerobic treatment use, and a variety of other factors.
- If this is a modeled dataset, what is the sample? UNCLEAR
- Output data is CO2 equivalencies in metric tonnes
- EPA data is accessible [here](https://www.epa.gov/statelocalenergy/download-state-inventory-and-projection-tool)
  MPCA data is accessible via the tool found [here](https://public.tableau.com/app/profile/mpca.data.services/viz/GHGemissioninventory/GHGsummarystory)
- We are currently using the default input values provided by the EPA, though will refine these based on our knowledge of local wastewater treatment in the future.
- Data is the 2020 statewide estimates downscaled to the county level by population percentage.
- This is the latest data available for both datasets (accessed January 2024).
- We are assuming EPA default input values are appropriate representations of the 11 county MSA region and that we can scale to county levels by population given that we are estimating municipal wastewater treatment.
- Which subject matter expert (SME) reviewed this data?
- We compare estimates from the MPCA and Met Council at the county level.

Be sure to add a citation of this dataset to the Zotero shared library.

### Data characteristics

- Were there any missing data? How did you handle missing data?
- Plots, tables, and description of data distribution
- Variance, Z-Score, quantiles
- Facet views by categorical variables

### Limitations

### Comparison with similar datasets

### Data dictionary

Table with detailed description of columns and definitions for each data table.

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
