## Wastewater
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_leaflet_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

# pull data here
epa_ww <- readRDS(file.path(here::here(), "_waste/data/epa_county_wastewater.RDS"))
mpca_ww <- readRDS(file.path(here::here(), "_waste/data-raw/wastewater/epa_wastewater.RDS"))
mc_ww <- readRDS(file.path(here::here(), "_waste/data-raw/wastewater/metc_wastewater.RDS"))
cprg_county <- readRDS(file.path(here::here(), "_meta/data/cprg_county.RDS"))
metc_service_area <- readRDS(file.path(here::here(), "_waste/data-raw/wastewater/metc_service_area.RDS"))


mc_ww_cty <- mc_ww %>%
  group_by(year, COUNTY_NAM, scope) %>%
  summarize(co2e = sum(value_emissions), .groups = "keep") %>%
  filter(year == 2021 & scope == "Scope 1")

hookaddcap()
```

We explored three sources for wastewater emissions: the EPA state inventory and projection tool, the Minnesota Pollution Control Agency (MPCA) emissions inventory, and Metropolitan Council Environmental Services emission estimates. Because only the EPA's estimate extends to Wisconsin and Minnesota, we are reporting those data for consistency, but explore all three below where they have geographic overlap.

The EPA estimate is a high-quality government dataset that models emissions using state population, biochemical oxygen demand, fraction of wastewater anaerobically digested, fraction of population on septic, estimated protein content, and biosolids used as fertilizer to estimate CH~4~ and N~2~O emissions [@usepaStateGreenhouseGas2024]. These inputs are modifiable and will be explored further in the CCAP. We apportioned the state level emission estimates to county levels by population percentage.

The EPA tool provides outputs in metric tons CO~2~e. We confirmed that the global warming potentials (GWP) used are from AR5 [@ipccClimateChange2014]. We used the most recent version of the tool released in January 2024, version 2024.1. 

We are currently using default input values provided by the EPA, though will refine these based on our knowledge of local wastewater treatment in the future.

You can access the EPA tool on the [EPA website](https://www.epa.gov/statelocalenergy/download-state-inventory-and-projection-tool) and the MPCA data through their [Tableau Dashboard](https://public.tableau.com/app/profile/mpca.data.services/viz/GHGemissioninventory/GHGsummarystory). You can also read full methodologies for MPCA [@claflinGreenhouseGasEmissions2023] and EPA [@usepaStateGreenhouseGas2024a] in their respective documents. 

### Data characteristics

There are many sources of uncertainty in these estimates due to multiple activity data inputs required for CH~4~ and N~2~O emission estimates. The EPA wastewater module does not offer guidance on quantifying these sources of uncertainty, only describing the potential sources themselves. We have not included any uncertainty estimates in our final values. 

### Limitations

- In order to have homogeneous comparisons among counties, we are relying on default EPA state inventory and projection tool data which may miss particularities about how wastewater is processed in our narrower county scope.
- We are assuming EPA default input values are appropriate representations of the 11 county MSA region and that we can scale to county levels by population given that we are estimating municipal wastewater treatment.

### Comparison with similar datasets

The Metropolitan Council, MPCA, and EPA source data leads to different estimates of emissions in the common seven county region. The Metropolitan Council does not process all wastewater in the seven county region. Additionally, they report CO~2~ emissions from combustion of stationary fuels for infrastructure used in directly processing wastewater, unlike EPA estimates which are CH~4~ and N~2~O emissions from municipal wastewater treatment of organic matter. 

```{r fig-mn-county-ww-emissions}
#| fig-cap: "County aggregated wastewater emissions by scope"
#| out-width: "95%"

metc_ww <- readRDS(file.path(here::here(), "_waste/data-raw/wastewater/metc_wastewater.RDS")) %>%
  filter(year == 2021)

metc_ww_summary <- metc_ww %>%
  group_by(COUNTY_NAM, year, scope) %>%
  summarize(
    value_emissions = sum(value_emissions),
    .groups = "keep"
  )

fig_mn_county_ww_emissions <- plot_ly(
  data = metc_ww_summary,
  x = ~ reorder(COUNTY_NAM, -value_emissions),
  y = ~value_emissions,
  source = "fig-mn-county-ww-emissions",
  color = ~scope,
  colors = c(
    "Scope 1" = colors$cdGreen,
    "Scope 2" = colors$councilBlue
  ),
  type = "bar",
  hovertemplate = ~ paste0(
    COUNTY_NAM, " County", "<br>",
    scope, "<br>",
    round(value_emissions / 1000, digits = 2), " thousand metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = "County wastewater emissions by scope",
    subtitle = "Metropolitan Council Environmental Services estimate",
    x_title = "County",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Scope"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "normal"
    )
  )

fig_mn_county_ww_emissions
```


```{r tbl-ww-comparison-scope}
#| tbl-cap: "Wastewater data sources and geographic scopes"
#| out-width: "95%"

tbl_ww_comparison_scope <- tribble(
  ~`Data source`, ~`Geographic scope`, # ~`Citation`,
  "EPA GHG Inventory and Projection Tool", "Minnesota and Wisconsin, scaled to county population", # "@usepaStateGreenhouseGas2024a",
  "MPCA GHG Inventory", "Minnesota, scaled to county population", # "@claflinGreenhouseGasEmissions2023",
  "Metropolitan Council Wastewater Emissions Estimate", "7-county Minnesota metro area" # "@"
) %>%
  gt() %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::fmt_markdown()

tbl_ww_comparison_scope
```


The Metropolitan Council area covers most of the 7-county Twin Cities area, but no areas in Wisconsin [@metcSewershedsMetershedsWWTP2020]. The Met Council environmental services division provided us with a city and county breakdown of wastewater emissions for their service area. 

<!-- We may want to move this map to the appendix? -->

```{r fig-map-ww-service-area}
#| fig-cap: "Metropoilitan Council wastewater service area"
#| out-width: "95%"
#| out-height: "500px"

fig_map_ww_service_area <- council_leaflet() %>%
  addPolygons(
    data = cprg_county %>%
      sf::st_transform(4326) %>%
      sf::st_union(),
    fill = FALSE,
    color = "gray",
    group = "CPRG area",
    popup = "CPRG county area"
  ) %>%
  addPolygons(
    data = metc_service_area %>%
      sf::st_simplify(dTolerance = 1),
    group = "MetC wastewater service area",
    weight = 1,
    fill = TRUE,
    fillColor = colors$esBlue,
    fillOpacity = 0.5,
    color = colors$councilBlue,
    popup = "Metropolitan Council wastewater service area"
  ) %>%
  leaflet::addLayersControl(
    overlayGroups = c(
      "CPRG area",
      "MetC wastewater service area"
    )
  )

fig_map_ww_service_area
```

Overall, the EPA estimate for each county proved highest, followed by the MPCA estimate and then Met Council estimate. We used the EPA estimate in our overall emissions estimate. 

```{r fig-wastewater-source-comparison}
#| fig-cap: "County wastewater emissions by data source"
#| out-width: "95%"

comp_2021 <- bind_rows(
  mc_ww_cty %>%
    mutate(county_name = COUNTY_NAM, source = "Met Council") %>%
    ungroup() %>%
    select(-c(year, scope, COUNTY_NAM)),
  mpca_ww %>%
    mutate(county_name = NAME, source = "MPCA", co2e = state_co2e) %>%
    ungroup() %>%
    select(-c(GEOID, NAME, state_co2e, epa_co2e)),
  epa_ww %>%
    mutate(county_name = NAME, source = "EPA", co2e = epa_co2e) %>%
    ungroup() %>%
    select(-c(GEOID, NAME, epa_co2e))
)

fig_wastewater_source_comparison <- plot_ly(
  data = comp_2021 %>%
    filter(county_name %in% mc_ww_cty$COUNTY_NAM),
  x = ~ reorder(county_name, -co2e),
  y = ~co2e,
  source = "fig-wastewater-source-comparison",
  type = "bar",
  hoverinfo = "text",
  color = ~source,
  colors = c(
    "EPA" = cprg_colors$cprg_da_yellow,
    "MPCA" = "#5D295F",
    "Met Council" = colors$councilBlue
  ),
  hovertemplate = ~ paste0(
    "County: ", county_name, "<br>",
    "Data source: ", source, "<br>",
    "Emissions: ", scales::comma(round(co2e)), " metric tons CO<sub>2</sub>e",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = "County wastewater emissions by data source",
    x_title = "County",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Data source",
    barmode = "identity"
  ) %>%
  layout(
    barmode = "identity"
  )

fig_wastewater_source_comparison
```

<!-- ### Data dictionary -->
<!-- Table with detailed description of columns and definitions for each data table. -->

Code reviewer: Liz Roten

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
