### Wisconsin Greenhouse Gas Emissions Inventory {#sec-wi-inv}

```{r}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
wi_emissions <- readRDS(file.path(here::here(), "_waste/data/solid_waste_WI_allyrs.RDS"))
wi_emissions_meta <- readRDS(file.path(here::here(), "_waste/data/solid_waste_WI_allyrs_meta.RDS"))
cprg_county <- readRDS(file.path(here::here(), "_meta/data/cprg_county.RDS"))
cprg_pop <- readRDS(file.path(here::here(), "_meta/data/cprg_population.RDS"))
county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

county_emissions_summary <- county_emissions %>%
  group_by(sector, category, source, data_source, factor_source) %>%
  summarize(
    value_emissions = sum(value_emissions),
    .groups = "keep"
  )
hookaddcap()
```


Waste emissions for Wisconsin counties were unavailable in the same detail as the MPCA data. Thus, we estimated total waste emissions based on statewide emissions estimates, allocated by population.

The most recent [Wisconsin Greenhouse Gas Emissions Inventory](https://widnr.widen.net/view/pdf/o9xmpot5x7/AM610.pdf?t.download=true) was done in 2021 by the Wisconsin Department of Natural Resources. Included in the solid waste data for this source are emissions from landfills and waste combustion, taking into account the emissions reduced by landfill gas collection for gas-to-energy use or flaring. This inventory does not, however, include emissions generated from composting or recycling.

The emissions for solid waste in this report were calculated using the [EPA's State Inventory and Projection Tool](https://www.epa.gov/statelocalenergy/state-inventory-and-projection-tool), a tool designed to help states calculate greenhouse gas emissions [@usepaStateGreenhouseGas2024a]. Default values provided by the tool were used except in the case of default Mixed Solid Waste population tonnage values, which were replaced by data from the Wisconsin state DNR Annual Waste Tonnage Report [@wisconsindnrGreenhouseGas2021].

For 2018, the Wisconsin DNR reported 2.2 million metric tons carbon dioxide equivalent (MMTCO~2~e) generated through landfilling and solid waste management.

In the process of analysis, this statewide estimate was apportioned to the county level based on county population data, as detailed in @sec-waste-methods.

### Limitations

- Since data reported directly from the counties was unavailable for Wisconsin, the solid waste data used here reflects a disaggregation of state-level data and may not be reflective of the specific mix of waste generated by Pierce and St. Croix counties.
- Data collected in Wisconsin's emissions inventory only represents waste disposed of in landfills or waste combustion facilities, and does not include organics. Composting data is unavailable for Wisconsin counties.

### Comparison to similar datasets

The US EPA completes yearly state-level estimates of emissions for each state, which combined sum to the totals reported in the US Greenhouse Gas Emissions Inventory. The data for these estimates and the US inventory can be explored at the [GHG Inventory Data Explorer](https://cfpub.epa.gov/ghgdata/inventoryexplorer/index.html). The EPA's total of landfill emissions for Wisconsin for 2018 was 2.450 MMTCO~2~e, not far off from the Wisconsin DNR's 2.2 MMTCO~2~e. The EPA's estimate for 2021 was 2.422 MMTCO~2~e. More details can be found in @sec-waste-inventory-comparison. 

Since the EPA completes an inventory for the entire US and its methods may not reflect the specific nuances of emissions in each state, we elected to use the data from the Wisconsin DNR for this inventory.

<!-- ### Data Dictionary -->


```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
