## Wastewater
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
# pull data here
epa_ww <- readRDS(file.path(here::here(), "_waste/data/epa_county_wastewater.RDS"))
mpca_ww <- readRDS(file.path(here::here(), "_waste/data-raw/wastewater/epa_wastewater.RDS"))
mc_ww <- readRDS(file.path(here::here(), "_waste/data-raw/wastewater/metc_wastewater.RDS"))


mc_ww_cty <- mc_ww %>%
  group_by(year, COUNTY_NAM, scope) %>%
  summarize(co2e = sum(emissions_metric_tons_co2e)) %>%
  filter(year == 2021 & scope == "Scope 1")

hookaddcap()
```

We currently have three sources for wastewater emissions: the EPA's state inventory tool, the MPCA's emissions inventory, and the Met Council's Environmental Services emission estimates. Because only the EPA's estimate extends to Wisconsin and Minnesota we are reporting those data for consistency, but explore all three below where they have geographic overlap. The EPA estimate uses state population, biochemical oxygen demand, fraction of wastewater anaerobically digested, fraction of population on septic, estimated protein content, and biosolids used as fertilizer to estimate CH4 and N2O emissions. These inputs are modifiable and should be explored further in the CCAP. We down-scaled the state level emission estimates to county levels by population percentage.

- Data is the highest quality rank (federal agency)
- Data is modeled by the agencies based on multiple parameters, including population, aerobic vs anaerobic treatment use, and a variety of other factors.
- If this is a modeled dataset, what is the sample? UNCLEAR
- Output data is CO2 equivalencies in metric tonnes
- EPA data is accessible [here](https://www.epa.gov/statelocalenergy/download-state-inventory-and-projection-tool)
  MPCA data is accessible via the tool found [here](https://public.tableau.com/app/profile/mpca.data.services/viz/GHGemissioninventory/GHGsummarystory)
- We are currently using the default input values provided by the EPA, though will refine these based on our knowledge of local wastewater treatment in the future.
- Data is the 2020 statewide estimates downscaled to the county level by population percentage.
- This is the latest data available for both datasets (accessed January 2024).
- We are assuming EPA default input values are appropriate representations of the 11 county MSA region and that we can scale to county levels by population given that we are estimating municipal wastewater treatment.
- Liz Roten is reviewing this data
- We compare estimates from the MPCA and Met Council at the county level.

Be sure to add a citation of this dataset to the Zotero shared library.

### Data characteristics

There are many sources of uncertainty in these estimates due to multiple activity data inputs required for CH4 and N2O emission estimates. The EPA Wastewater module does not offer guidance on quantifying these sources of uncertainty, only describing the potential sources themselves.

### Limitations

In order to have homogenous comparisons among counties, we are relying on default EPA State Inventory Tool data which may miss particularities about how wastewater is processed in our narrower county scope.

### Comparison with similar datasets

The Metropolitan Council, MPCA, and EPA source data leads to different estimates of emissions in the common seven county region. Metropolitan Council does not process all wastewater in the seven county region. Additionally they report CO2 emissions from combustion of stationary fuels for infrastructure used in directly processing wastewater, unlike EPA estimates which are CH4 and N2O emissions from municipal wastewater treatment of organic matter.

```{r fig-wastewater-source-comparison}
#| fig-cap: "Total waste generated by county"
#| out-width: "95%"

# is this too many categories. should i switch to side-by-side bar plot

comp_2021 <- bind_rows(
  mc_ww_cty %>%
    mutate(county_name = COUNTY_NAM, source = "Met Council") %>%
    ungroup() %>%
    select(-c(year, scope, COUNTY_NAM)),
  mpca_ww %>%
    mutate(county_name = NAME, source = "MPCA", co2e = state_co2e) %>%
    ungroup() %>%
    select(-c(GEOID, NAME, state_co2e, epa_co2e)),
  epa_ww %>%
    mutate(county_name = NAME, source = "EPA", co2e = epa_co2e) %>%
    ungroup() %>%
    select(-c(GEOID, NAME, epa_co2e))
)

comp_2021 %>%
  filter(county_name %in% mc_ww_cty$COUNTY_NAM) %>%
  plot_ly(
    x = ~ reorder(county_name, -co2e),
    y = ~co2e,
    source = opts_current$get()$label,
    type = "bar",
    hoverinfo = "text",
    color = ~source,
    hovertemplate = ~ paste0(
      "County: ", county_name, "<br>",
      "Data source: ", source, "<br>",
      "Emissions: ", round(co2e), " metric tonnes CO2e",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "Wastewater emissions per county by source",
    x_title = "County",
    y_title = "CO2e emissions (metric tonnes)",
    legend_title = "Data source",
    barmode = "identity"
  ) %>%
  layout(
    barmode = "identity"
  )
```

### Data dictionary

Table with detailed description of columns and definitions for each data table.

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
