# Data sources and validation

```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
hookaddcap()
```

## MPCA Data

The Minnesota Pollution Control Agency collects an annual inventory, [SCORE](https://public.tableau.com/app/profile/mpca.data.services/viz/SCOREOverview/1991-2021SCORE), of solid waste, recycling and organics as reported by counties across the state. For this project, we used MPCA data for the nine Minnesota counties for the year 2021.

Data source description, type

- Quality rank (See @tbl-quality-rank)
Highest quality
- How, when, and why was the data collected?
- If this is a modeled dataset, what is the sample?
not modeled - directly reported
- What is the raw unit of measurement? 
- How was this data accessed? Include any relevant links/citations, code, or downloads.
- What data cleaning or wrangling was completed? How did you test these processes and outputs?
- What is the geographic and temporal scope? Did you complete any aggregation?
annual, county-level
- What version is the data? Were there other versions available? If so, why did you choose this version?
- What assumptions are made when we use this dataset?
- Which subject matter expert (SME) reviewed this data?
- Describe testing used to verify data

Be sure to add a citation of this dataset to the Zotero shared library.

### Data characteristics

- Were there any missing data? How did you handle missing data?
- Plots, tables, and description of data distribution
- Variance, Z-Score, quantiles
- Facet views by categorical variables


```{r echo=FALSE}
saveCap("waste-data")
```
