## WI Waste Data

```{r}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
# pull data here
wi_emissions <- readRDS(file.path(here::here(), "_waste/data/wi_emissions.RDS"))
wi_emissions_meta <- readRDS(file.path(here::here(), "_waste/data/wi_emissions_meta.RDS"))
# PULL IN DATA FROM TRANSPORTATION TO USE:
cprg_county <- readRDS(file.path(here::here(), "R/data/cprg_county.RDS"))
cprg_pop <- readRDS(file.path(here::here(), "R/data/cprg_population.RDS"))
hookaddcap()
```


Waste emissions for Wisconsin counties were unavailable in the same detail as the MPCA data. Thus, we estimated total waste emissions based on statewide emissions estimates, allocated by population.

## WI Greenhouse Gas Emissions Inventory

The most recent [Wisconsin Greenhouse Gas Emissions Inventory](https://widnr.widen.net/view/pdf/o9xmpot5x7/AM610.pdf?t.download=true) was done in 2018 by the Wisconsin Department of Natural Resources. Included in the solid waste data for this source are emissions from landfills and waste combustion, taking into account the emissions reduced by waste-to-energy technology or flaring [further research needed]. This inventory does not, however, include emissions generated from composting or recycling.

For 2018, the Wisconsin DNR reported 2.2 Million Metric Tons Carbon Dioxide Equivalent (MMTCO2e) generated through landfilling and solid waste management.

The US EPA also completes yearly state-level estimates of emissions for each state, which combined sum to the totals reported in the US Greenhouse Gas Emissions Inventory. The data for these estimates and the US inventory can be explored at the [GHG Inventory Data Explorer](https://cfpub.epa.gov/ghgdata/inventoryexplorer/index.html). The EPA's total of landfill emissions for Wisconsin for 2018 was 2.450 MMTCO2e, not far off from the Wisconsin DNR's 2.2 MMTCO2e. The EPA's estimate for 2021 was 2.422 MMTCO2e.

Since the EPA completes an inventory for the entire US and its methods may not reflect the specific nuances of emissions in each state, we elected to use the data from the Wisconsin DNR for this inventory.

- Quality rank (See @tbl-quality-rank)
Highest quality - state-level source
- How, when, and why was the data collected?
- If this is a modeled dataset, what is the sample?
not modeled - directly reported
- What is the raw unit of measurement? 
- How was this data accessed? Include any relevant links/citations, code, or downloads.
- What data cleaning or wrangling was completed? How did you test these processes and outputs?
- What is the geographic and temporal scope? Did you complete any aggregation?
- What version is the data? Were there other versions available? If so, why did you choose this version?
- What assumptions are made when we use this dataset?
- Which subject matter expert (SME) reviewed this data?
- Describe testing used to verify data


```{r echo=FALSE}
saveCap("data-wi-emissions")
```
