# Deliverables (Landfill)
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/plot_county_emissions.R"))
hookaddcap()
```

Local inventory of GHG emissions from landfills with documentation of the following QC activities: 

(1) narrative report describing data sources and QC measures for data acquisition steps, 

(2) description of methodology and QC measures for validated proper implementation of methodology, and

(3) documentation of QAPP implementation.

(4) listing of emissions reductions options are present with documentation of rationale for each option.


## Quality Control {#sec-waste-qc-procedure}


1.	Comparison of (a) independent local inventory versus (b) landfill data from FLIGHT. Use a table similar to the table below to assess precision and bias of the local inventory versus QC estimates. Additionally, precision and bias calculations will be in accordance with the [EPA’s Data Assessment Statistical Calculator (DASC) Tool](https://www.epa.gov/sites/default/files/2020-10/dasc_11_3_17.xls) with the community’s estimate taken as the measured value and the LGGIT [@usepaLocalGreenhouseGas2017] value taken as the audit value.

2.	When comparing any two datasets, ensure that the units of measure are converted to a consistent basis prior to making the comparison.

3.	Ensure the GWPs used for the local estimate and independent estimate are on the same basis. 

4.	Ensure data are appropriate for intended use, data are complete and representative and current, data sources are documented, analytical methods are appropriate, and calculations are accurate. Include any QC findings and reconciliation.

5.	Review by TL or senior technical reviewer—analytical methods and results are explained clearly, technical terms are defined, conclusions are reasonable based on information presented, and level of technical detail is appropriate)

6.	Editor review—writing is clear, free of grammatical and typing errors.


```{r tbl-waste-qc}
#| tbl-cap: "Landfill emissions comparison between local estimate and FLIGHT."
#| out-width: "95%"

tbl_waste_qc <-
  tibble::tribble(
    ~"Solid Waste (Landfills)", ~"Initial Local Estimate (Metric Tons CO~2~e)", ~"FLIGHT Data (Metric Tons CO~2~e)",
    "", "", "",
    "", "", "",
    "", "", "",
    "", "", ""
  ) %>%
  gt() %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::fmt_markdown() %>%
  gt::as_raw_html()

tbl_waste_qc
```


```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
