# Solid Waste 
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
cprg_population <- readRDS(file.path(here::here(), "_meta/data/cprg_population.RDS"))
epa_waste_inventory <- readRDS(file.path(here::here(), "_waste/data-raw/epa_solidwaste_inventory.RDS"))

federal_totals <- epa_waste_inventory %>%
  filter(source == "Total") %>%
  select(
    geog_name,
    source,
    emissions_metric_tons_co2e,
    data_source
  )

cprg_ctu <- readRDS(file.path(here::here(), "_meta/data/cprg_ctu.RDS"))
county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

county_emissions_summary <- county_emissions %>%
  group_by(sector, category, source, data_source, factor_source) %>%
  summarize(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e),
    .groups = "keep"
  )

sector_percentages <- county_emissions_summary %>%
  group_by(sector) %>%
  summarize(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)) %>%
  arrange(-emissions_metric_tons_co2e) %>%
  janitor::adorn_percentages(denominator = "col")


waste_percentages <- county_emissions_summary %>%
  filter(sector == "Waste") %>%
  group_by(category, source) %>%
  summarize(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e), .groups = "keep") %>%
  arrange(-emissions_metric_tons_co2e) %>%
  janitor::adorn_percentages(denominator = "col")
hookaddcap()
```

## Introduction

(Overview of emissions sources)

(Key processes leading to emissions)
In 2021, waste activities over the entire United States generated emissions of 169.2 MMTCO~2~e, or 2.7 percent of total U.S. greenhouse gas emissions [@usepaInventoryGreenhouseGas2023]. Solid waste emissions primarily consist of CH~4~, emitted in large part by landfills and organics composting, but also include CO~2~ and N~2~O generation. In the metropolitan area, waste incineration accounts for a significant portion of solid waste management and therefore emissions. 

## Results

Emissions from waste is the smallest of the three sectors in the Twin Cities MSA. Waste generated `r round(sum(county_emissions$emissions_metric_tons_co2e)/1000000 * sector_percentages[[2]][3], digits = 2)` MMtCO~2~e of emissions in the Twin Cities MSA in 2021. Solid waste, including landfills, recycling, and organics, generates the largest share of emissions in the waste sector, with municipal wastewater treatment comprising a smaller share of waste emissions. 

Waste emissions in Minnesota have declined nearly 40% since 2005, due to a variety of factors including gas capture technologies and aging waste in open landfills [@mpcaClimateChangeTrends2023]. Waste emissions in Wisconsin show no significant change from 2005 to 2018, possibly due to increased waste generation offsetting reductions from gas capture [@wisconsindnrGreenhouseGas2021]. The Metropolitan Council, which oversees wastewater services for the majority of the urbanized extent of the metropolitan region, continues to work to reduce wastewater emissions across its 9 wastewater treatment plants.

```{r fig-waste-wastewater-county-emissions}
#| fig-cap: "Solid waste and wastewater county emissions"
#| out-width: "95%"
fig_waste_wastewater_county_emissions <-
  plot_county_sector_emissions(county_emissions,
    .sector = "Waste",
    .plotly_source = "fig-waste-wastewater-county-emissions"
  )

fig_waste_wastewater_county_emissions$x$layoutAttrs[[1]]$title$text <- "Solid waste and wastewater emissions"

fig_waste_wastewater_county_emissions
```


```{r fig-waste-county-emissions-by-category}
#| fig-cap: "Solid waste and wastewater county emissions by category"
#| out-width: "95%"

fig_waste_county_emissions_by_category <- plot_county_emissions(
  county_emissions = county_emissions,
  .sector = "Waste",
  .plotly_source = "fig-waste-county-emissions-by-category"
)

fig_waste_county_emissions_by_category$x$layoutAttrs[[1]]$title$text <- "Solid waste and wastewater emissions"

fig_waste_county_emissions_by_category
```


### Solid waste

```{r solid-waste-percentages, include=FALSE}
solid_waste_percentages <- county_emissions %>%
  filter(category == "Solid waste") %>%
  group_by(category, source) %>%
  summarize(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)) %>%
  arrange(-emissions_metric_tons_co2e) %>%
  janitor::adorn_percentages(denominator = "col")


total_solid_waste <-
  county_emissions %>%
  filter(category == "Solid waste") %>%
  ungroup() %>%
  summarise(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)) %>%
  extract2("emissions_metric_tons_co2e")
```

Solid waste generated `r round(total_solid_waste/1000000, digits = 2)` MMtCO~2~e of emissions in the Twin Cities MSA in 2021. Of that total, `r scales::percent(solid_waste_percentages$emissions_metric_tons_co2e[1], accuracy = 0.1)` of emissions came from `r str_to_lower(solid_waste_percentages$source[1])`, `r scales::percent(solid_waste_percentages$emissions_metric_tons_co2e[2], accuracy = 0.1)` from `r str_to_lower(solid_waste_percentages$source[2])` facilities, and the remaining `r scales::percent(sum(solid_waste_percentages$emissions_metric_tons_co2e[3:4]), accuracy = 0.1)` from `r str_to_lower(solid_waste_percentages$source[3])` and `r str_to_lower(solid_waste_percentages$source[4])`.  

```{r fig-solid-waste-emissions}
#| fig-cap: "2021 solid waste emissions"
#| out-width: "95%"
fig_solid_waste_emissions <-
  plot_ly(
    type = "bar",
    source = "fig-solid-waste-emissions",
    data = county_emissions %>%
      filter(category == "Solid waste"),
    y = ~ reorder(geog_name, emissions_metric_tons_co2e),
    x = ~emissions_metric_tons_co2e,
    color = ~category,
    colors = unlist(category_colors),
    marker = list(
      line = list(
        color = colors$councilBlue
      )
    ),
    hovertemplate = ~ paste0(
      geog_name, " County", "<br>",
      category, ", ", source, "<br>",
      round(emissions_metric_tons_co2e / 1000, digits = 0), " thousand metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "2021 solid waste emissions",
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Category"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    yaxis = list(categoryorder = "total ascending")
  )
fig_solid_waste_emissions
```

Greenhouse gas emissions from solid waste are dominated by the landfill sector. In Hennepin, Ramsey, and Washington counties, municipal centers where a significant portion of waste is incinerated, waste-to-energy or incineration makes up a large fraction of emissions as well. 

```{r fig-solid-waste-sector-emissions-by-category}
#| fig-cap: "2021 solid waste emissions by category"
#| out-width: "95%"

fig_solid_waste_sector_emissions_by_category <-
  plot_ly(
    type = "bar",
    source = "fig-solid-waste-sector-emissions-by-category",
    data = county_emissions %>%
      filter(category == "Solid waste"),
    y = ~ reorder(geog_name, emissions_metric_tons_co2e),
    x = ~emissions_metric_tons_co2e,
    color = ~source,
    colors = unlist(source_colors),
    hovertemplate = ~ paste0(
      geog_name, " County", "<br>",
      category, "<br>",
      round(emissions_metric_tons_co2e / 1000, digits = 0), " thousand metric tons CO<sub>2</sub>e", "<br>",
      "<extra></extra>"
    )
  ) %>%
  plotly_layout(
    main_title = "2021 solid waste emissions by category",
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Category"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    ),
    yaxis = list(categoryorder = "total ascending")
  )
fig_solid_waste_sector_emissions_by_category
```



```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
