# Methods and data sources  

```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

cprg_census_county_population <- readRDS(file.path(here::here(), "_meta/data/census_county_population.RDS")) %>%
  filter(cprg_area == TRUE) %>%
  mutate(population_year = as.numeric(population_year))
county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

onroad_emissions <- readRDS(file.path(here::here(), "_transportation/data/onroad_emissions.RDS")) %>%
  rename(value_emissions = emissions_metric_tons_co2e)

county_inventory_forecast_emissions <- readRDS(file.path(here::here(), "_transportation/data/county_inventory_forecast_emissions.RDS")) %>%
  rowwise() %>%
  mutate(rounded_tons = round_emissions_metric_tons_co2e(emissions_metric_tons_co2e))


transportation_max_year <- onroad_emissions %>%
  filter(emissions_year == max(emissions_year)) %>%
  extract2("emissions_year") %>%
  unique()

transportation_min_year <- onroad_emissions %>%
  filter(emissions_year == min(emissions_year)) %>%
  extract2("emissions_year") %>%
  unique()

modes_included <-
  stringr::str_split(onroad_emissions$vehicle_types_included, pattern = ", ", simplify = TRUE)[, 1] %>%
  unique() %>%
  stringr::str_to_lower()

mndot_vmt_ctu_gap_filled <- readRDS(file.path(here::here(), "_transportation/data/mndot_vmt_ctu_gap_filled.RDS"))

hookaddcap()
```

## Methodological framework

Transportation emissions are available for `r listify(modes_included)` using EPA data sources. We use a geographic, or territorial accounting method, aligning with Scope 1 in the Greenhouse Gas Protocol. Geographic methods account for any transportation emissions taking place within a geographic boundary, regardless of origin or destination [@fongGlobalProtocolCommunityScale2021].

Geographic emissions are essential for quantifying air pollution experienced by people living in the area, but they do not give information on the logistic decisions of individuals.

<!-- Met Council researchers chose to use an alternative activity-induced method to provide the most actionable information for jurisdictions within our region in both the recent PCAP and the 2018 Metro Climate Stats inventory. -->

<!-- TODO include summary table with geographic level, year, data source -->

```{r tbl-inventory-forecast-sources-county}
#| tbl-cap: "County emissions inventory and forecast data sources"
#| out-width: "95%"

tbl_inv_forecast_sources <- county_inventory_forecast_emissions %>%
  ungroup() %>%
  select(emissions_year, data_source, county_name) %>%
  unique() %>%
  arrange(-emissions_year) %>%
  group_by(data_source) %>%
  summarize(
    Years = paste0(
      min(emissions_year), " - ", unique(max(emissions_year), min(emissions_year))
    ),
    Counties = paste0(unique(county_name), collapse = ", ")
  ) %>%
  mutate(Years = case_when(
    Years == "2050 - 2050" ~ "2050",
    Years == "2020 - 2020" ~ "2020",
    TRUE ~ Years
  )) %>%
  arrange(Years)

gt(tbl_inv_forecast_sources) %>%
  gt::cols_label("data_source" = c("Data source")) %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::cols_align(align = "left")
```


The Met Council also completed city-level VMT estimates for cities within the 7-county metro area. 

```{r tbl-inventory-forecast-sources-city}
#| tbl-cap: "City VMT history and forecast data sources"
#| out-width: "95%"

tbl_ctu_modeled_vmt <- mndot_vmt_ctu_gap_filled %>%
  select(inventory_year, final_vmt_source, ctu_name_full_county) %>%
  unique() %>%
  arrange(ctu_name_full_county) %>%
  group_by(final_vmt_source) %>%
  summarize(
    Years = paste0(
      min(inventory_year), " - ", unique(max(inventory_year), min(inventory_year))
    ),
    Cities = paste0(unique(ctu_name_full_county), collapse = "; ")
  ) %>%
  mutate(
    Years = case_when(
      Years == "2050 - 2050" ~ "2050",
      TRUE ~ Years
    ),
    Cities = case_when(
      final_vmt_source == "Regional Travel Demand Model" ~ "All",
      TRUE ~ Cities
    )
  )

gt(tbl_ctu_modeled_vmt) %>%
  gt::cols_label("final_vmt_source" = c("Data source")) %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::cols_align(align = "left")
```

All data processing for transportation is compiled using [_run_all_transportation.R](data-raw/_run_all_transportation.R), ensuring all disparate datasets and sequential processing requirements are correct.  

## Data sources

<!-- {{< include data_aviation.qmd >}} -->

{{< include data_epa_emissions.qmd >}}

<!-- {{< include data_streetlight.qmd >}} -->

{{< include data_state_dot.qmd >}}

<!-- {{< include data_epa_moves.qmd >}} -->

<!-- {{< include data_tbi.qmd >}} -->

{{< include data_regional_travel_demand_model.qmd >}}

{{< include data_original_analysis.qmd >}}


## Limitations {#sec-transportation-data-limitations}

- Geographic accounting methods
  - Geographic accounting methods do not account for the decisions or travel behavior of individuals within the geographic boundaries. Within the Twin Cities region, this method will show high emissions per capita in low population areas with significant vehicle traffic, such as a small town with a major freeway. The residents of the small town are not responsible for the emissions of vehicles passing through their town without stopping, but they are subject to the air pollution associated with those trips. Additionally, the city governing body cannot expect to reduce emissions from freeway traffic, as the road is out of their jurisdiction.
  
<!-- - We will take these limitations into account and plan to mitigate wherever possible for CTU-level inventories and forecasts. -->

## Validation

### Correlation with related data

We would expect counties with higher population to have higher emissions and VMT.

```{r fig-emissions-population-transportation}
#| fig-cap: "County population and transportation emissions"
#| out-width: "95%"

onroad_emissions_population <- onroad_emissions %>%
  group_by(emissions_year, county_name, geoid, data_source) %>%
  summarize(
    value_emissions = sum(value_emissions),
    .groups = "keep"
  ) %>%
  rowwise() %>%
  left_join(cprg_census_county_population,
    by = c("geoid", "county_name",
      "emissions_year" = "population_year"
    )
  )

fig_emissions_pop_transportation <- plot_ly(
  data = onroad_emissions_population %>%
    filter(emissions_year == transportation_max_year),
  source = "fig-emissions-population-transportation",
  x = ~population,
  y = ~value_emissions,
  type = "scatter",
  mode = "markers",
  hovertemplate = ~ paste0(
    county_name, " County", "<br>",
    "Population: ", scales::comma(population), "<br>",
    "Emissions: ", round(value_emissions * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  ),
  marker = list(
    color = colors$councilBlue,
    size = 18,
    line = list(
      color = "lightgray",
      width = 2
    )
  )
) %>%
  plotly_layout(
    main_title = "County population and emissions",
    subtitle = paste0(transportation_max_year, " ACS 5-year population estimates. Strong positive correlation"),
    x_title = "Total population",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "",
    legend = list(orientation = "h")
  )

fig_emissions_pop_transportation
```

<!-- ### Comparison with other inventories -->

<!-- #### National Emissions Inventory -->

<!-- The National Emissions Inventory (NEI) is a comprehensive and detailed estimate of air emissions of criteria pollutants, criteria precursors, and hazardous air pollutants from air emissions sources. The county-level GHG emissions included in the NEI for this category are calculated by running the MOVES model with State-, Local-, and Tribal-submitted activity data and EPA-developed activity inputs based on data from FHWA and other sources [@usepa2020NationalEmissions2023].  -->

<!-- NEI data were pulled using the [EnviroFacts](https://enviro.epa.gov/) API and processed in R scripts: [epa_nei.R](../_meta/data-raw/epa_nei.R) and [epa_nei_envirofacts.R](data-raw/epa_nei_envirofacts.R) -->

```{r read-process-nei}
epa_nei <- readRDS(file.path(here::here(), "_transportation/data/epa_nei_envirofacts.RDS"))


vmt_emissions <- readRDS(file.path(here::here(), "_transportation/data/county_vmt_emissions.RDS"))
vmt_emissions_all <- vmt_emissions %>%
  group_by(zone) %>%
  summarise(
    vmt_total = sum(vmt_total),
    vmt_same = sum(vmt_same),
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)
  )


nei_emissions_all <- epa_nei %>%
  filter(nei_inventory_year == 2020) %>%
  group_by(zone = county_name) %>%
  summarize(
    value_emissions = sum(emissions_metric_tons_co2e)
  )

nei_cprg_combine <- left_join(
  vmt_emissions_all %>%
    mutate(cprg_emissions = emissions_metric_tons_co2e) %>%
    select(-emissions_metric_tons_co2e),
  nei_emissions_all %>%
    mutate(nei_emissions = value_emissions) %>%
    select(-value_emissions)
)
```

<!-- We expect that the NEI data will show higher emissions, because it is based on overall activity, not explicit origin-destination travel.  -->

<!-- ```{r fig-emissions-council-nei} -->
<!-- #| fig-cap: "Council county emissions estimate correlation with NEI estimate" -->
<!-- #| out-width: "95%" -->

<!-- nei_cprg_combine <- left_join( -->
<!--   vmt_emissions_all %>% -->
<!--     mutate(cprg_emissions = value_emissions) %>% -->
<!--     select(-value_emissions), -->
<!--   nei_emissions_all %>% -->
<!--     mutate(nei_emissions = value_emissions) %>% -->
<!--     select(-value_emissions) -->
<!-- ) -->

<!-- fig_nei_council_emissions <- plot_ly( -->
<!--   data = nei_cprg_combine, -->
<!--   source = "fig-nei-council-emissions", -->
<!--   x = ~nei_emissions, -->
<!--   y = ~cprg_emissions, -->
<!--   type = "scatter", -->
<!--   mode = "markers", -->
<!--   hovertemplate = ~ paste0( -->
<!--     zone, " County", "<br>", -->
<!--     "NEI estimate: ", round(nei_emissions * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>", -->
<!--     "Met Council estimate: ", round(cprg_emissions * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>", -->
<!--     "<extra></extra>" -->
<!--   ), -->
<!--   marker = list( -->
<!--     color = colors$councilBlue, -->
<!--     size = 18, -->
<!--     line = list( -->
<!--       color = "lightgray", -->
<!--       width = 2 -->
<!--     ) -->
<!--   ) -->
<!-- ) %>% -->
<!--   plotly_layout( -->
<!--     main_title = "Calculated county emissions correlation with NEI emissions", -->
<!--     # subtitle = "2021 ACS 5-year population estimates. Strong positive correlation", -->
<!--     x_title = "NEI estimate (metric tons CO<sub>2</sub>e)", -->
<!--     y_title = "Met Council estimate (metric tons CO<sub>2</sub>e)", -->
<!--     legend_title = "", -->
<!--     legend = list(orientation = "h") -->
<!--   ) -->

<!-- fig_nei_council_emissions -->
<!-- ``` -->


<!-- ```{r fig-activity-vs-nei} -->
<!-- #| fig-cap: "County estimated emissions, NEI and calculated" -->
<!-- #| out-width: "95%" -->

<!-- fig_activity_vs_nei <- plot_ly( -->
<!--   data = vmt_emissions_all, -->
<!--   source = "fig-activity-vs-nei", -->
<!--   x = ~ reorder(zone, -value_emissions), -->
<!--   y = ~value_emissions, -->
<!--   type = "scatter", -->
<!--   mode = "markers", -->
<!--   name = "2021 Met Council estimate", -->
<!--   hovertemplate = ~ paste0( -->
<!--     zone, " County<br>", -->
<!--     "Council estimate: ", round(value_emissions * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>", -->
<!--     "<extra></extra>" -->
<!--   ), -->
<!--   opacity = 0.8, -->
<!--   marker = list( -->
<!--     color = colors$councilBlue, -->
<!--     size = 18, -->
<!--     line = list( -->
<!--       color = "lightgray", -->
<!--       width = 2 -->
<!--     ) -->
<!--   ) -->
<!-- ) %>% -->
<!--   add_trace( -->
<!--     data = nei_emissions_all, -->
<!--     name = "2020 NEI on-road estimate", -->
<!--     marker = list( -->
<!--       color = cprg_colors$cprg_da_yellow, -->
<!--       size = 18, -->
<!--       line = list( -->
<!--         color = "lightgray", -->
<!--         width = 2 -->
<!--       ) -->
<!--     ), -->
<!--     hovertemplate = ~ paste0( -->
<!--       zone, " County<br>", -->
<!--       "NEI estimate: ", round(value_emissions * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>", -->
<!--       "<extra></extra>" -->
<!--     ) -->
<!--   ) %>% -->
<!--   plotly_layout( -->
<!--     main_title = "County estimated vehicle emissions by data source", -->
<!--     subtitle = "", # "2020 NEI, all on-road emissions", -->
<!--     x_title = "County", -->
<!--     y_title = "Metric tons CO<sub>2</sub>e", -->
<!--     legend_title = "Data source" -->
<!--   ) -->

<!-- fig_activity_vs_nei -->
<!-- ``` -->




<!-- #### Local Greenhouse Gas Inventory Tool (LGGIT) {#sec-epa-lggit-transportation} -->

<!-- EPA’s Local Greenhouse Gas Inventory Tool (LGGIT) was developed to help communities across the United States to evaluate their greenhouse gas emissions. We used the LGGIT to validate our calculations and datasets [@usepaLocalGreenhouseGas2017]. -->

<!-- We used the LGGIT Excel workbook tool with the following inputs in the "Mobile-Data" sheet. Several calculations and assumptions were made -->

<!-- - Passenger VMT was split between gasoline and diesel powered vehicles based on regional fleet composition in the Travel Behavior Inventory (TBI). -->
<!-- - Passenger vehicle age was determined from the median vehicle model year in the TBI. See @sec-regional-fleet for more detail. -->
<!-- - Fuel consumption was calculated by dividing VMT by average vehicle miles per gallon (MPG), specific to vehicle type and fuel type. -->
<!-- - Fuel efficiency data were taken from the LGGIT tool and verified with the EPA Emission Factors Hub (2021 edition) [@usepaGHGEmissionFactors2021].   -->
<!-- - All medium-duty VMT were classified as "Light Truck", while all heavy-duty VMT were classified as "Heavy-Duty Vehicle".   -->
<!-- - Commercial trucks were assumed to be year 2007, based on a 2018 report that found the average age of trucks in the US to be 14.2 years [@brusseauAgingTrucksCreate2019] (`2021 - 14 = 2007`).    -->


<!-- LGGIT entries and resulting output tables were generated and processed in [epa_lggit.R](data-raw/epa_lggit.R).  -->


<!-- ```{r tbl-lggit-entries} -->
<!-- #| tbl-cap: "Mobile data entered into LGGIT" -->
<!-- #| out-width: "95%" -->
<!-- lggit_vmt_entries <- readRDS(file.path(here::here(), "_transportation/data-raw/epa/lggit_vmt_entries.RDS")) -->

<!-- tbl_lggit_entries <- lggit_vmt_entries %>% -->
<!--   gt() %>% -->
<!--   gt::opt_row_striping() %>% -->
<!--   fmt_number(8:9, decimals = 0) %>% -->
<!--   tab_style( -->
<!--     style = cell_text(weight = "bold"), -->
<!--     locations = cells_column_labels() -->
<!--   ) %>% -->
<!--   gt::as_raw_html() -->

<!-- tbl_lggit_entries -->
<!-- ``` -->


```{r lggit-calculations}
cprg_total_emiss <- vmt_emissions_all %>%
  ungroup() %>%
  summarise(value_emissions = sum(emissions_metric_tons_co2e)) %>%
  magrittr::extract2("value_emissions") %>%
  scales::comma()

lggit_totals <- readRDS(file.path(here::here(), "_transportation/data-raw/epa/lggit_totals.RDS"))

lggit_total_emiss <- (lggit_totals$total %>% sum()) %>% scales::comma()
```

<!-- The results from the LGGIT tool estimate a regional emissions total of **`r lggit_total_emiss` metric tons CO~2~e**, while our calculations estimate **`r cprg_total_emiss` metric tons CO~2~e**. These values are sufficiently close, and may be explained by differences between nationwide vehicle emission rates and fuel efficiency and region-specific values from EPA MOVES. Additionally, LGGIT uses global warming potential (GWP) values from the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment Report (AR5) [@ipccClimateChange2014], while our calculations use slightly higher values from the 6th Assessment Report (AR6) [@ipccAR62021]. See @sec-epa-moves for a more detailed comparison. -->

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```

{{< pagebreak >}}

