# Mobile combustion  
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
hookaddcap()
```
## Introduction

Transportation is consistently the highest source of emissions in the state (MPCA report). 

## Methods

The general workflow for transportation emissions is to find vehicle miles traveled for a given geography, 

### Vehicle Miles Traveled      

To find the estimate number of vehicle miles traveled (VMT), we used the estimated number of vehicles and the average trip length in miles for all origin - destination pairs. VMT is calculated as follows:

$$VMT = trips\times length$$

where 

$$trips = \text{number of trips}\\length = \text{average trip length in miles}$$

## Results



```{r echo=FALSE}
saveCap("trans")
```
