# Mobile combustion  
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)


knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r, include=FALSE}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))

# vmt_emissions <- readRDS(file.path(here::here(), "_transportation/data/county_vmt_emissions.RDS"))
# vmt_emissions_all <- vmt_emissions %>%
#   group_by(zone) %>%
#   summarise(
#     vmt_total = sum(vmt_total),
#     vmt_same = sum(vmt_same),
#     emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)
#   )


onroad_emissions <- readRDS(file.path(here::here(), "_transportation/data/onroad_emissions.RDS"))

onroad_county_summary <- onroad_emissions %>% 
  group_by(emissions_year, county_name, geoid, data_source) %>% 
  summarize(emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e),
            .groups = "keep") %>% 
  rowwise() %>% 
  mutate(rounded_tons = round_emissions_metric_tons_co2e(emissions_metric_tons_co2e))

county_emissions <- readRDS(file.path(here::here(), "_meta/data/cprg_county_emissions.RDS"))

county_emissions_summary <- county_emissions %>%
  group_by(sector, category, source, data_source, factor_source) %>%
  summarize(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e),
    .groups = "keep"
  )

transportation_max_year <-onroad_emissions %>%
  filter(emissions_year == max(emissions_year)) %>%
  extract2("year") %>%
  unique()


transportation_min_year <- onroad_emissions %>%
  filter(emissions_year == min(emissions_year)) %>%
  extract2("year") %>%
  unique()

modes_included <-
  stringr::str_split(onroad_emissions$vehicle_types_included, pattern = ", ", simplify = TRUE)[, 1] %>%
  unique() %>%
  stringr::str_to_lower()

hookaddcap()
```

## Introduction

Transportation emissions, when grouped together, are one of the largest sources of greenhouse gas emissions nationally, statewide, and locally. [more information here on transportation emissions, what is included/not included, ]

This is the transportation section for estimating on-road emissions. Modes included are `r listify(modes_included)`. 

Note that this does not include freight rail, light rail, and other transportation emissions.

## `r transportation_max_year` emissions

```{r fig-transportation-county-emissions}
#| fig-cap: `{r} paste0(transportation_max_year, " onroad transportation emissions by county")`
#| out-width: "95%"
fig_transportation_county_emissions <- plot_county_sector_emissions(
  county_emissions %>%
    filter(year == transportation_max_year),
  .sector = "Transportation",
  .plotly_source = "fig-transportation-county-emissions"
)
fig_transportation_county_emissions
```

We can also view this data broken out by vehicle weight in @fig-transportation-county-emissions-by-weight. 


```{r fig-transportation-county-emissions-by-weight}
#| fig-cap: `{r} paste0(transportation_max_year, " annual emissions by vehicle weight")`
#| out-width: "95%"

fig_transportation_county_emissions_by_weight <- plot_ly(
  type = "bar",
  source = "fig-transportation-county-emissions-by-weight",
  data = onroad_emissions %>% 
    filter(emissions_year == transportation_max_year),
  y = ~reorder(county_name, emissions_metric_tons_co2e),
  x = ~emissions_metric_tons_co2e,
  color = ~vehicle_weight_label,
  colors = unlist(vehicle_weight_colors),
  hovertemplate = ~ paste0(
    county_name, " County", "<br>",
    vehicle_weight_label, "<br>",
    round(emissions_metric_tons_co2e * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = paste0(transportation_max_year, " annual emissions by vehicle weight"),
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Vehicle weight"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    )
  )
fig_transportation_county_emissions_by_weight
```


```{r fig-transportation-county-emissions-by-fuel-type}
#| fig-cap: `{r} paste0(transportation_max_year, " annual emissions transportation by fuel type")`
#| out-width: "95%"

fig_transportation_county_emissions_by_fuel <- plot_ly(
  type = "bar",
  source = "fig-transportation-county-emissions-by-fuel-type",
  data = onroad_emissions %>% 
    filter(emissions_year == transportation_max_year),
  y = ~reorder(county_name, emissions_metric_tons_co2e),
  x = ~emissions_metric_tons_co2e,
  color = ~vehicle_fuel_label,
  colors = unlist(fuel_type_colors),
  hovertemplate = ~ paste0(
    county_name, " County", "<br>",
    vehicle_fuel_label, "<br>",
    round(emissions_metric_tons_co2e * 1e-6, digits = 2), " million metric tons CO<sub>2</sub>e", "<br>",
    "<extra></extra>"
  )
) %>%
  plotly_layout(
    main_title = paste0(transportation_max_year, " annual transportation emissions by fuel type"),
    subtitle = "",
    y_title = "County",
    x_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "Fuel type"
  ) %>%
  layout(
    barmode = "stack",
    legend = list(
      traceorder = "reversed"
    )
  )
fig_transportation_county_emissions_by_fuel
```


## Historical emissions

```{r fig-transportation-time-series-county}
#| fig-cap: `{r} paste0(("Transportation emissions, ", transportation_min_year, " - ", transportation_max_year))`
#| out-width: "95%"


fig_transportation_time_series_county <- onroad_county_summary %>% 
  group_by(county_name) %>% 
  plot_ly(
    source = "fig-transportation-time-series-county",
    x = ~emissions_year,
    y = ~emissions_metric_tons_co2e,
    color = ~county_name,
    colors = "Dark2",
    type = "scatter",
    mode = "lines+markers",
    marker = list(size = 10),
    hoverinfo = "text",
    hovertext = ~paste0(
      county_name, "County, ", emissions_year, " ", "<br>",
      rounded_tons
    )
  ) %>% 
  plotly_layout(
    main_title = paste0("Transportation emissions, ", transportation_min_year, " - ", transportation_max_year),
    subtitle = "",
    x_title = "Year",
    y_title = "Metric tons CO<sub>2</sub>e",
    legend_title = "County"
  ) 

fig_transportation_time_series_county
```

