## Travel Behavior Inventory {#sec-tbi}
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
veh21 <- readRDS(file.path(here::here(), "_transportation/data-raw/tbi/veh21.RDS")) %>%
  mutate(
    fuel_orig = forcats::fct_rev(fuel_orig),
    fuel = forcats::fct_rev(fuel)
  )
trip21 <- readRDS(file.path(here::here(), "_transportation/data-raw/tbi/trip21.RDS"))
regional_trip_length_avg <- readRDS(file.path(here::here(), "_transportation/data-raw/tbi/tbi_regional_trip_length_avg.RDS"))
tbi_od_ordered_trip_length <- readRDS(file.path(here(), "_transportation/data-raw/tbi/tbi_od_ordered_trip_length.RDS"))
tbi_vehicle_fuel_age <- readRDS(file.path(here::here(), "_transportation/data-raw/tbi/tbi_vehicle_fuel_age.RDS"))
tbi_home_county <- readRDS(file.path(here::here(), "_transportation/data-raw/tbi/tbi_home_county.RDS"))


hookaddcap()
```


The Metropolitan Council Travel Behavior Inventory (TBI) is a bi-annual household survey around 7,500 families in the 7-county Twin Cities metro and three neighboring Wisconsin counties. Information on people, households, trips, and vehicles are collected [@metcTravelBehaviorInventory2021]. This survey was active in the field from May 22, 2021 to February 5, 2023.  

Data were imported directly from a Council-maintained [public GitHub repository](https://github.com/Metropolitan-Council/metc.tbi.helper). The calculations below were verified from other Council analysis projects using the same dataset. Exactly 7,745 households with homes in any of the 11 CPRG counties were included in this subset of the TBI dataset.


```{r tbl-tbi-hh-county-dist}
#| tbl-cap: "2021 TBI household survey geographic distribution"
#| out-width: "95%"

tbi_home_county %>%
  select(
    "Household county" = cprg_hh_county,
    "Estimated number of households" = est_n,
    "Estimated number of households standard error" = est_n_se,
    "Estimated percentage of households" = est_pct,
    "Estimated percentage of all households standard error" = est_pct_se,
    "Sample size" = n
  ) %>%
  gt::gt() %>%
  gt::fmt_number(2:3, decimals = 0) %>%
  gt::fmt_percent(4:5) %>%
  gt::fmt_number(6, decimals = 0) %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::as_raw_html()
```


### Regional fleet characteristics {#sec-regional-fleet}

We used 2021 TBI data to determine the regional average vehicle age and distribution of diesel and gasoline passenger vehicles of households in the CPRG study area. 

Vehicles were classified into broad fuel categories - diesel and gas + all other fuels (including gasoline, electric, flex-fuel, hybrid, and plug-in hybrid) to best match the average miles per gallon table specifications in the EPA Local Greenhouse Gas Inventory Tool (LGGIT). The resulting value is on par with recent statistics from the Bureau of Transportation Statistics (BTS), which calculates the average passenger vehicle age in 2021 to be 12.1 years [@btsAverageAgeAutomobiles2023]. 

TBI data were cleaned to only include vehicles with complete data and model year 1980 or later. Vehicles with a fuel type "Other (e.g., natural gas, bio-diesel)" were removed due to low sample size. 

```{r tbl-vehicle-fuel-age}
#| tbl-cap: "Median vehicle age and proportion of all regional vehicles by grouped fuel type"
#| out-width: "95%"
tbi_vehicle_fuel_age %>%
  select(
    "Fuel type" = fuel,
    "Median vehicle year" = year_median,
    "Median vehicle year standard error" = year_median_se,
    "Estimated number of vehicles" = est_n,
    "Estimated number of vehicles standard error" = est_n_se,
    "Estimated percentage of all vehicles" = est_pct,
    "Estimated percentage of all vehicles standard error" = est_pct_se,
    "Sample size" = n
  ) %>%
  gt::gt() %>%
  gt::fmt_number(3:6, decimals = 0) %>%
  gt::fmt_number(8, decimals = 0) %>%
  gt::fmt_percent(6:7) %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::as_raw_html()
```

```{r fig-regional-fuel-veh-age}
#| fig-cap: "Regional vehicle fleet model year by fuel type"
#| out-width: "95%"
#| out-height: "500px"

fuel_dist_pl <- plot_ly(
  type = "box",
  data = veh21,
  x = ~year,
  y = ~fuel_orig,
  name = "Original fuel types"
) %>%
  plotly_layout()


fuel_grp <- plot_ly(
  type = "box",
  data = veh21,
  x = ~year,
  y = ~fuel,
  name = "Grouped fuels"
) %>%
  plotly_layout()

subplot(
  fuel_grp,
  fuel_dist_pl,
  shareX = TRUE,
  heights = c(0.2, 0.8),
  nrows = 2
) %>%
  plotly_layout(
    main_title = "Regional fleet vehicle year by fuel type",
    subtitle = "2021 Travel Behavior Inventory",
    x_title = "Vehicle model year",
    y_title = ""
  ) %>%
  layout(
    legend = list(orientation = "h")
  )
```

### Average trip distance between counties

The average trip distance for the entire region is `r round(regional_trip_length_avg$mean_trip_dist, 2)` miles (standard error `r round(regional_trip_length_avg$mean_trip_dist_se, 2)`), based on a sample of `r scales::comma(regional_trip_length_avg$n)`  trips.   

Trips with distance over 720 miles (the equivalent of 12 hours of driving at 60 miles per hour) were removed. Only Minnesota CPRG counties were available for analysis.

We used the TBI to validate StreetLight's average trip length. See @sec-trip-length-validation for more detail. 

```{r tbl-reg-avg-trip-length}
#| tbl-cap: "Regional mean trip distance. 2021 TBI."
#| out-width: "95%"

regional_trip_length_avg %>%
  select(
    "Mean trip distance (miles)" = mean_trip_dist,
    "Mean trip distance standard error" = mean_trip_dist_se,
    "Estimated number of trips" = estimate_n,
    "Estimated number of trips standard error" = estimate_n_se,
    "Distance variance" = var,
    "Sample size" = n
  ) %>%
  gt() %>%
  fmt_number(c(3, 4, 5, 6),
    decimals = 0
  ) %>%
  fmt_number(1:2,
    decimals = 2
  ) %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::as_raw_html()
```

Origin-destination pairs with fewer than 30 observed trips were removed. 

:::{.scrolling}

```{r tbl-od-trip-length}
#| tbl-cap: "Mean trip distance by origin-destination county"
#| out-width: "95%"
tbi_od_ordered_trip_length %>%
  arrange(-estimate_n) %>%
  select(
    "Origin-Destination pair" = origin_dest_county_pair,
    "Sample size" = n,
    "Mean trip distance (miles)" = mean_trip_dist,
    "Mean trip distance standard error" = mean_trip_dist_se,
    "Estimated number of trips" = estimate_n,
    "Estimated number of trips standard error" = estimate_n_se,
    "Distance variance" = var
  ) %>%
  gt() %>%
  fmt_number(c(2, 5, 6),
    decimals = 0
  ) %>%
  fmt_number(c(3:4,7),
    decimals = 2
  ) %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::as_raw_html()
```

:::

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
