## StreetLight Data
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
hookaddcap()
```

Introduction text 
Data source description, type

- Qualtiy rank (See @tbl-quality-rank)
- How, when, and why was the data collected?
- If this is a modeled dataset, what is the sample?
- What is the raw unit of measurement? 
- How was this data accessed? Include any relevant links/citations, code, or downloads.
- What data cleaning or wrangling was completed? How did you test these processes and outputs?
- What is the geographic and temporal scope? Did you complete any aggregation?
- What version is the data? Were there other versions available? If so, why did you choose this version?
- What assumptions are made when we use this dataset?
- Which subject matter expert (SME) reviewed this data?
- Describe testing used to verify data

Be sure to add a citation of this dataset to the references.bib

### Data characteristics

- Were there any missing data? How did you handle missing data?
- Plots, tables, and description of data distribution
- Variance, Z-Score, quantiles
- Facet views by categorical variables

```{r fig-scatter-stl}
#| fig-cap: "A scatter plot"
shinipsum::random_ggplotly(type = "point") %>%
  councilR::plotly_layout(
    main_title = "Scatter"
  )
```


```{r fig-boxplot-stl}
#| fig-cap: "A boxplot"
shinipsum::random_ggplotly(type = "boxplot") %>%
  councilR::plotly_layout(
    main_title = "Boxplot"
  )
```


```{r fig-density-stl}
#| fig-cap: "A density plot"
shinipsum::random_ggplotly(type = "density") %>%
  councilR::plotly_layout(
    main_title = "Density"
  )
```


### Limitations

### Comparison with similar datasets

### Data dictionary

Table with detailed description of columns and definitions for each data table.

```{r echo=FALSE}
saveCap("stl-data")
```
