#### National Emissions Inventory {#sec-epa-nei}
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))


epa_onroad_source_set <- readRDS(file.path(here::here(), "_transportation/data/epa_onroad_source_set.RDS"))
pollutant_key <- readRDS(file.path(here::here(), "_transportation/data/pollutant_key.RDS"))
pollutant_key_meta <- readRDS(file.path(here::here(), "_transportation/data/pollutant_key_meta.RDS"))
epa_onroad_emissions_compile <- readRDS(file.path(here::here(), "_transportation/data/epa_onroad_emissions_compile.RDS"))

epa_onroad_year_source <- epa_onroad_emissions_compile %>%
  mutate(emissions_year = as.character(emissions_year)) %>%
  select(emissions_year, data_source) %>%
  unique()

hookaddcap()
```


The [National Emissions Inventory](https://www.epa.gov/air-emissions-inventories/national-emissions-inventory-nei) (NEI) is a comprehensive and detailed estimate of air emissions of criteria pollutants, criteria precursors, and hazardous air pollutants from air emissions sources. The county-level GHG emissions included in the NEI for this category are calculated by running the MOVES model with State-, Local-, and Tribal-submitted activity data and EPA-developed activity inputs based on data from FHWA and other sources [@usepa2020NationalEmissions2023]. 

NEI data were pulled using the [EnviroFacts](https://enviro.epa.gov/) API and processed in R scripts: [epa_nei.R](../_meta/data-raw/epa_nei.R) and [epa_nei_envirofacts.R](data-raw/epa_nei_envirofacts.R).

NEI SMOKE FF10 data are processed in [epa_nei_smoke_ff.R](data-raw/epa_nei_smoke_ff.R). 

NEI on-road regional summaries are processed in [epa_nei_onroad_emissions.R](data-raw/epa_nei_onroad_emissions.R). 

Ultimately, NEI data used in the Metropolitan Council inventory were compiled from SMOKE FF10 for year `r epa_onroad_year_source %>% filter(data_source == "National Emissions Inventory") %>% extract2("emissions_year")`. 

```{r read-nei}
epa_nei <- readRDS(file.path(here::here(), "_transportation/data/epa_nei_envirofacts.RDS"))

nei_emissions_all <- epa_nei %>%
  filter(nei_inventory_year == 2020) %>%
  group_by(zone = county_name) %>%
  summarize(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)
  )
```


##### Verification and validation {#sec-nei-verif-valid}

NEI data were cross-verified by comparing county level emissions totals compiled from NEI EnviroFacts, NEI data summaries by region, and compiled SMOKE FF10. 

[epa_verify_nei_envirofacts_smoke.R](data-raw/epa_verify_nei_envirofacts_smoke.R) found that data compiled from SMOKE FF10 and regional summaries aligned exactly for year 2020 and closely for other years. Similarly, data compiled from EnviroFacts also aligned closely with SMOKE FF10 and regional summaries. 

Data published on the EPA website are subject to change at any time. Every effort was taken to align versions, model runs, and other opportunities for differentiation. 


<!-- ##### Activity data {#sec-epa-nei-activity} -->

<!-- TODO Compare underlying activity data with data from state DOTs -->

<!-- ##### Emissions factors -->

<!-- #### Limitations  -->



```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
