## EPA MOVES {#sec-epa-moves}
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
epa_moves <- readRDS(file.path(here::here(), "_transportation/data/epa_moves.RDS"))
epa_moves_meta <- readRDS(file.path(here::here(), "_transportation/data/epa_moves_meta.RDS"))
hookaddcap()
```

Emissions rates for our region were calculated using the EPA's [Motor Vehicle Emissions Simulator (MOVES4)](https://www.epa.gov/moves) [@usepaMotorVehicleEmission2023]. MOVES4 calculates emissions factors using Council's regional travel demand model, Minnesota Department of Vehicle Services county vehicle registration data, and the Minnesota Pollution Control Agency (MPCA) vehicle age distribution. Each of these inputs helps the model estimate the characteristics of vehicles on the road in our region. The model takes into account differences in fuel economy (miles per gallon) depending on a vehicle's age and size, as well as its fuel intake (diesel or gasoline). The results are specific to the conditions of our region, and so are more accurate than national averages. 

<!-- TODO fill in these values -->
MOVES is a high-quality government modeling system with strong data inputs and outputs. We requested this data from our MOVES coordinator, Dennis Farmer, who ran the model and shared the resulting table with us. The model run covers [exact geographic scope here] for years [exact years here], using MOVES version [exact version here]. 

The resulting table provides grams of CO~2~, CH~4~, and N~2~O per vehicle mile traveled. We imported, processed, and saved the data in an R script, [epa_moves.R](data-raw/epa_moves.R). CO~2~ equivalence (CO~2~e) values are derived using global warming potential (GWP) values. See @sec-gwp for more details.

```{r clean-emiss-rates}
epa_moves_clean <- epa_moves %>%
  rowwise() %>%
  mutate(
    `Kilograms CH4 per mile` = ch4 / 1000,
    `Kilograms N2O per mile` = n2o / 1000,
    `Kilograms CO2 per mile` = co2 / 1000,
    `Kilograms CO2e per mile` = co2_co2_equivalent / 1000
  ) %>%
  mutate(
    `Grams CH₄ per mile` = `Kilograms CH4 per mile` * 1000,
    `Grams N₂O per mile` = `Kilograms N2O per mile` * 1000,
    `Grams CO₂ per mile` = `Kilograms CO2 per mile` * 1000,
    `Grams CO₂e per mile` = `Kilograms CO2e per mile` * 1000
  ) %>%
  select(
    `Vehicle weight` = vehicle_weight,
    `MOVES year` = moves_year,
    starts_with("grams")
  )

lggit_kg_emissions_per_mile <- readRDS(file.path(here::here(), "_transportation/data-raw/epa/lggit_kg_emissions_per_mile.RDS"))

lggit_clean <- lggit_kg_emissions_per_mile %>%
  select(
    `Vehicle type` = `Vehicle Type`,
    `Vehicle model year` = `Vehicle Year`,
    `Fuel type` = `Fuel Type`,
    `Average miles per gallon`,
    `Kilograms CH4 per mile`,
    `Kilograms N2O per mile`,
    `Kilograms CO2 per mile`,
    `Kilograms CO2e per mile`
  ) %>%
  mutate(
    `Grams CH₄ per mile` = `Kilograms CH4 per mile` * 1000,
    `Grams N₂O per mile` = `Kilograms N2O per mile` * 1000,
    `Grams CO₂ per mile` = `Kilograms CO2 per mile` * 1000,
    `Grams CO₂e per mile` = `Kilograms CO2e per mile` * 1000
  ) %>%
  select(
    `Vehicle type`,
    `Vehicle model year`,
    `Fuel type`,
    starts_with("grams")
  )
```


```{r tbl-epa-moves-values}
#| tbl-cap: "Emissions in grams per mile traveled by vehicle weight in the Twin Cities region. EPA MOVES4."
#| out-width: "95%"
tbl_epa_moves_values <- epa_moves_clean %>%
  gt() %>%
  gt::opt_row_striping() %>%
  fmt_number(3:6, decimals = 2) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::as_raw_html()
tbl_epa_moves_values
```
### Comparsion with EPA GHG Emissions Hub (2021)

For comparison, we pulled the emissions per mile rates from the Local Greenhouse Gas Inventory Tool (LGGIT), which align with the 2021 EPA GHG Emission Hub [@usepaGHGEmissionFactors2021]. 

```{r tbl-lggit-emissions-per-mile}
#| tbl-cap: "Grams of emissions per mile by vehicle type and fuel type. EPA GHG Emission Hub (2021)"
#| out-width: "95%"

tbl_lggit_emissions_per_mile <- lggit_clean %>%
  gt() %>%
  fmt_number(4:7, decimals = 2) %>%
  gt::opt_row_striping() %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  gt::as_raw_html()

tbl_lggit_emissions_per_mile
```

```{r lggit-mean-passenger}
# find weighted mean LGGIT based on
# the proportion of Gasoline/Diesel vehicles in the region
lggit_pass <- lggit_clean %>%
  mutate(
    data_source = "EPA GHG Hub",
    `Vehicle weight` = case_when(
      `Vehicle type` == "Passenger Car" ~ "Passenger",
      `Vehicle type` == "Light Truck" ~ "Medium",
      `Vehicle type` == "Heavy-Duty Vehicle" ~ "Heavy"
    )
  ) %>%
  select(
    -`Grams CH₄ per mile`, -`Grams N₂O per mile`,
    -`Grams CO₂ per mile`
  ) %>%
  pivot_longer(
    cols = 4,
    names_to = "ghg_factor"
  ) %>%
  filter(`Vehicle weight` == "Passenger") %>%
  mutate(weight_value = c(0.98, 0.02)) %>%
  group_by(`Vehicle weight`, data_source, ghg_factor) %>%
  summarize(value = weighted.mean(value, weight_value), .groups = "keep")

# combine MOVES and GHG hub tables
epa_plus_moves <- bind_rows(
  epa_moves_clean %>%
    mutate(data_source = "MOVES"),
  lggit_clean %>%
    filter(`Vehicle type` != "Passenger Car") %>%
    mutate(
      data_source = "EPA GHG Hub",
      `Vehicle weight` = case_when(
        `Vehicle type` == "Passenger Car" ~ "Passenger",
        `Vehicle type` == "Light Truck" ~ "Medium",
        `Vehicle type` == "Heavy-Duty Vehicle" ~ "Heavy"
      )
    )
) %>%
  select(
    -`Grams CH₄ per mile`, -`Grams N₂O per mile`,
    -`Grams CO₂ per mile`
  ) %>%
  pivot_longer(
    cols = 3,
    names_to = "ghg_factor"
  ) %>%
  bind_rows(lggit_pass)

# create wide version for segments
epa_moves_wide <- epa_plus_moves %>%
  select(
    -ghg_factor, -`Vehicle model year`,
    -`Vehicle type`, -`MOVES year`, -`Fuel type`
  ) %>%
  pivot_wider(
    names_from = data_source,
    values_from = value
  )
```

To directly compare overall passenger emissions rates, we applied a weighted average to the EPA GHG Hub emissions rates for passenger vehicles according to the regional fleet gasoline/diesel distribution (98% gasoline, 2% diesel). Learn more about the regional fleet in @sec-regional-fleet.  

```{r fig-moves-ghg-hub-comparison}
#| fig-cap: "MOVES4 and GHG Emissions Hub per mile emission rates by vehicle weight"
#| out-width: "95%"
#| out-height: "300px"

fig_moves_ghg_hub_comparison <- plot_ly(
  data = epa_plus_moves %>%
    group_by(`Vehicle weight`),
  source = opts_current$get()$label,
  y = ~`Vehicle weight`,
  x = ~value,
  color = ~data_source,
  type = "scatter",
  mode = "markers",
  marker = list(
    size = 10
  ),
  hovertemplate = ~ paste0(
    "Vehicle weight: ", `Vehicle weight`, "<br>",
    "Data source: ", data_source, "<br>",
    scales::comma(value), " grams CO<sub>2</sub>e per mile",
    "<extra></extra>"
  )
) %>%
  add_segments(
    inherit = FALSE,
    data = epa_moves_wide,
    y = ~`Vehicle weight`,
    yend = ~`Vehicle weight`,
    x = ~MOVES,
    xend = ~`EPA GHG Hub`,
    showlegend = FALSE
  ) %>%
  plotly_layout(
    main_title = "MOVES4 and GHG Emissions Hub per mile emission rates",
    x_title = "Grams CO<sub>2</sub>e per mile"
  ) %>%
  layout(barmode = "identity")

fig_moves_ghg_hub_comparison
```

### Limitations

- This edition of MOVES is fairly old relative to our estimation year (2021).
- We are not breaking out vehicles by fuel type; instead, we are aggregating based on the regional fleet. This may result in more inaccuracies.
- MOVES4 only accounts for vehicles that are registered in the 7-county metro area, so does not account for vehicles on regional roads, but registered elsewhere. However, the traffic generated from those vehicles is detected in the regional travel demand model. 

Subject matter expert reviewer: ________

<!-- ### Data dictionary -->

<!-- ```{r tbl-epa-moves-meta} -->
<!-- #| tbl-cap: "EPA MOVES metadata" -->
<!-- #| out-width: "95%" -->
<!-- epa_moves_meta %>% -->
<!--   gt() %>% -->
<!--   gt::opt_row_striping() %>% -->
<!--   tab_style( -->
<!--     style = cell_text(weight = "bold"), -->
<!--     locations = cells_column_labels() -->
<!--   ) %>% -->
<!--   gt::fmt_markdown() %>% -->
<!--   gt::as_raw_html() -->
<!-- ``` -->



```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
