### Air Emissions Modeling Platforms {#sec-epa-air-emissions-modeling}
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
hookaddcap()


epa_onroad_source_set <- readRDS(file.path(here::here(), "_transportation/data/epa_onroad_source_set.RDS"))
pollutant_key <- readRDS(file.path(here::here(), "_transportation/data/pollutant_key.RDS"))
pollutant_key_meta <- readRDS(file.path(here::here(), "_transportation/data/epa_onroad_source_set.RDS"))
epa_onroad_emissions_compile <- readRDS(file.path(here::here(), "_transportation/data/epa_onroad_emissions_compile.RDS"))


epa_onroad_year_source <- epa_onroad_emissions_compile %>%
  mutate(emissions_year = as.character(emissions_year)) %>%
  select(emissions_year, data_source) %>%
  unique()

aem_years <- epa_onroad_emissions_compile %>%
  filter(data_source == "Air Emissions Modeling") %>%
  select(data_source, emissions_year, interpolation) %>%
  unique()
```

The EPA continually works on emissions inventories for various projects. 

Air Emissions Modeling data are available for several years, but only years `r aem_years$emissions_year %>% listify()` are used in the final inventory.

Both the 2021 and 2022 estimates are based on the 2020 NEI [@usepa2022v2EmissionsModeling2025, @usepa2021EmissionsModeling2024]. 

Air Emissions Modeling SMOKE FF10 data are processed in [epa_air_emissions_modeling_onroad.R](data-raw/epa_air_emissions_modeling_onroad.R). 

#### Verification and validation {#sec-epa-air-emis-verif-valid}

Air Emissions Modeling data are only available from a single consistent website, and so verification across locations was not necessary. 

<!-- #### Activity data {#sec-epa-air-emis-activity-data} -->

<!-- TODO Compare underlying activity data with data from state DOTs -->

<!-- #### Emissions factors -->

#### Limitations {#sec-epa-air-emis-mod-limit}

In addition to limitations described in @sec-epa-emissions-limitations, Air Emissions Modeling has its own set of limitations.

- Air Emissions Modeling datasets are in active development and subject to change.   

```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
