## National Emissions Inventory {#sec-epa-nei}
```{r}
source(file.path(here::here(), "R/_load_pkgs.R"))
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_plotting_helpers.R"))
hookaddcap()
```


The [National Emissions Inventory](https://www.epa.gov/air-emissions-inventories/national-emissions-inventory-nei) (NEI) is a comprehensive and detailed estimate of air emissions of criteria pollutants, criteria precursors, and hazardous air pollutants from air emissions sources. The county-level GHG emissions included in the NEI for this category are calculated by running the MOVES model with State-, Local-, and Tribal-submitted activity data and EPA-developed activity inputs based on data from FHWA and other sources [@usepa2020NationalEmissions2023]. 

NEI data for aircraft and locomotives in our region are unavailable prior to 2017. 

We are assuming that all light-duty vehicles are passenger vehicles, and all heavy-duty are commercial. 

NEI data were pulled using the [EnviroFacts](https://enviro.epa.gov/) API and processed in R scripts: [epa_nei.R](../_meta/data-raw/epa_nei.R) and [epa_nei_envirofacts.R](data-raw/epa_nei_envirofacts.R)

```{r read-nei}
epa_nei <- readRDS(file.path(here::here(), "_transportation/data/epa_nei_envirofacts.RDS"))

nei_emissions_all <- epa_nei %>%
  filter(nei_inventory_year == 2020) %>%
  group_by(zone = county_name) %>%
  summarize(
    emissions_metric_tons_co2e = sum(emissions_metric_tons_co2e)
  )
```


## Comparison of underlying activity data

Compare underlying activity data with data from state DOTs

### Vehicle miles traveled

## Limitations 



```{r echo=FALSE}
caption_index <- readRDS(file.path(here::here(), "caption_index.RDS"))
caption_index <- as.character(as.numeric(caption_index) + 1) %>% stringr::str_pad(width = 2, side = "left", pad = "0")
saveCap(paste0("cap-", caption_index))
saveRDS(caption_index, file.path(here::here(), "caption_index.RDS"))
```
{{< pagebreak >}}
