# Deliverables
```{r  include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.pos = "H",
  out.width = "100%",
  dpi = 300
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
```{r}
source(file.path(here::here(), "R/_quarto_helpers.R"))
source(file.path(here::here(), "R/_load_pkgs.R"))
hookaddcap()
```

Local inventory of GHG emissions from mobile sources with documentation of the following QC activities: 

(1) narrative report describing data sources and QC measures for data acquisition steps, 
(1) description of methodology and QC measures for validated proper implementation of methodology, and
(1) documentation of QAPP implementation.
(1) listing of emissions reductions options are present with documentation of rationale for each option.


## Quality Control {#sec-transportation-qc-procedure}


1. Comparison of local estimate of average miles traveled per year and average miles per gallon (by vehicle type) versus state and national averages. 

1. For any values used in local inventory that differ from the state average MPY or the national average MPG by more than 10% the community will provide an explanation of why local factors may differ from state or national averages. Additionally, precision and bias calculations will be in accordance with the [EPA’s Data Assessment Statistical Calculator (DASC) Tool](https://www.epa.gov/sites/default/files/2020-10/dasc_11_3_17.xls) with the community’s estimate taken as the measured value and the LGGIT [@usepaLocalGreenhouseGas2017] value taken as the audit value.


1. Ensure the GWPs used for the local estimate and the LGGIT [@usepaLocalGreenhouseGas2017] estimate are on the same basis. The LGGIT tool uses AR5 GWP (e.g., methane GWP = 28).

1. Review by TL or senior technical reviewer — analytical methods / results are explained clearly, technical terms are defined, conclusions are reasonable based on information presented, and level of technical detail is appropriate.

1. Editor review — verify or remediate draft deliverables to ensure clear, error-free writing.

```{r tbl-transport-qc-mpy}
#| tbl-cap: "Transportation average miles per year, local and quality control comparison by vehicle and fuel types."
tibble::tribble(
  ~"Vehicle Type", ~"Local Avg Miles/yr", ~"QC Avg Miles/yr",
  "Passenger Car (Gasoline)", "", "",
  "Passenger Truck (Gasoline)", "", "",
  "Heavy-duty (Gasoline)", "", "",
  "Motorcycle (Gasoline)", "", "",
  "Passenger Car (Diesel)", "", "",
  "Passenger Truck (Diesel)", "", "",
  "Heavy-duty (Diesel)", "", ""
) %>%
  knitr::kable()
```
```{r tbl-transport-qc-mpg}
#| tbl-cap: "Transportation average miles per gallon, local and quality control comparison by vehicle and fuel types."
tibble::tribble(
  ~"Vehicle Type", ~"Local Avg Miles/gal", ~"QC Avg Miles/gal",
  "Passenger Car (Gasoline)", "", 24.1,
  "Passenger Truck (Gasoline)", "", 18.5,
  "Heavy-duty (Gasoline)", "", 10.1,
  "Motorcycle (Gasoline)", "", 50,
  "Passenger Car (Diesel)", "", 32.4,
  "Passenger Truck (Diesel)", "", 22.1,
  "Heavy-duty (Diesel)", "", 13.0
) %>%
  knitr::kable()
```

```{r echo=FALSE}
saveCap("qc-transportation")
```
